---
title: "1154 RP Diff_analysis"
output: html_notebook
---

---
title: "1154 RP Differential Analysis"
output: html_notebook
---

---
title: "RP Differential Analysis"
output: html_notebook
---

# 1. Importing Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggrepel)
library(ggsci)
library(pheatmap)
library(RColorBrewer)
library(readxl)
library(UpSetR)
source('functions_cdis_diff.R')
```

# 2. Import data


Because differential analysis includes the calculations of means, the data to be used in this section is the normalized untransformed data

```{r set_path, message=FALSE, warning=FALSE}

# Load compounds table and add a column that has names for identified compounds and FeatureID for the rest (for plotting puRPouses)
project_dir <- getwd()
project_name <- 'RP_LCMS_Germ'
figures_dir <- file.path(project_dir, paste0(project_name, '_output_figures'))
tables_dir <- file.path(project_dir,  paste0(project_name, '_output_tables'))
# If the flag was setted before, nothing needs to be changed here, the correct file will be automatically used

compounds_table_file <- file.path(tables_dir, 'compounds_table.csv')
compounds_table <- read_csv(compounds_table_file)

# Import metadata and fix names
metadata_file <- file.path(tables_dir, 'fixed_metadata.csv')
metadata <- read_csv(metadata_file)
metadata<- subset(metadata, Fungi=='1154')
# set path variables
project_dir <- getwd()
project_name <- '1154_RP_LCMS_Germ'
figures_dir <- file.path(project_dir, paste0(project_name, '_output_figures'))
tables_dir <- file.path(project_dir,  paste0(project_name, '_output_tables'))


norm_auc_table_file <-file.path(tables_dir,'normalized_untransformed_auc_table.csv')
# Load auc_table

norm.matrix <- read_csv(norm_auc_table_file)
norm.matrix <- column_to_rownames(norm.matrix, var = '...1')



```



# 3. Calculate ratios and log2fold-change

## 3.1 Get samples per each treatment

```{r}

# Get the average AUC per each of the treatments
## Get samples per treatment

# Get the average AUC per each of the treatments
## Get samples per treatment

BOCU.samples <- get_samples(metadata, Treatment = 'Plant', value = 'BOCU')
CTRL.samples <- get_samples(metadata, Treatment = 'Plant', value = 'CTRL')
ERIN.samples <- get_samples(metadata, Treatment = 'Plant', value = 'ERIN')
LEDU.samples <- get_samples(metadata, Treatment = 'Plant', value = 'LEDU')
ERLE.samples <- get_samples(metadata, Treatment = 'Plant', value = 'ERLE')

```

## 3.2 Get table of differentially expressed features per each comparison

The following function will calculate ratio, log2FC, p values and adjusted pvalues. If no replicates are available for EACH treatment please use the get_diff_table_no_pval() function

```{r}

# The following function will calculate ratio, log2FC, p values and adjusted pvalues. If no replicates are available for EACH treatment
# please use the get_diff_table_no_pval() function

BOCU_to_CTRL.diff_table <- get_diff_table(norm.matrix, treatment.sample_list = BOCU.samples, control.sample_list = CTRL.samples)
ERLE_to_CTRL.diff_table <- get_diff_table(norm.matrix, treatment.sample_list = ERLE.samples, control.sample_list = CTRL.samples)
ERIN_to_CTRL.diff_table <- get_diff_table(norm.matrix, treatment.sample_list = ERIN.samples, control.sample_list = CTRL.samples)
LEDU_to_CTRL.diff_table <- get_diff_table(norm.matrix, treatment.sample_list = LEDU.samples, control.sample_list = CTRL.samples)
```

Merge differentially expressed features with the annotation

```{r}

# Create dataframes for the up and downregulated metabolites at each time point and merge them with the compound information

BOCU_to_CTRL.diff_table <- compounds_table %>% 
  select(-contains('Annotation Source:'), - contains('Results'), -SampleID, -AUC,  -Fungi, - Plant) %>% 
  right_join(BOCU_to_CTRL.diff_table, by = 'FeatureID') %>% 
  distinct() %>% 
  filter(!is.nan(log2FC)) %>% 
  mutate(Comment = case_when(log2FC == Inf ~ 'Not present in control',
                             log2FC == -Inf ~ 'Only present in control',
                             log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

BOCU_to_CTRL.diff_table$Comment <- factor(BOCU_to_CTRL.diff_table$Comment, 
                               levels = c('Only present in control', 'Downregulated', 'Upregulated', 'Not present in control' ))

table_file <- file.path(tables_dir, 'Diff_expressed_BOCU.csv')
write_csv(BOCU_to_CTRL.diff_table, table_file )

ERLE_to_CTRL.diff_table<- compounds_table %>% 
  select(-contains('Annotation Source:'), - contains('Results'), -SampleID, -AUC, -Plant, - Fungi) %>%
  right_join(ERLE_to_CTRL.diff_table, by = 'FeatureID') %>% 
  distinct() %>% 
  filter(!is.nan(log2FC)) %>% 
  mutate(Comment = case_when(log2FC == Inf ~ 'Not present in control',
                             log2FC == -Inf ~ 'Only present in control',
                             log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

ERLE_to_CTRL.diff_table$Comment <- factor(ERLE_to_CTRL.diff_table$Comment, 
                               levels = c('Only present in control', 'Downregulated', 'Upregulated', 'Not present in control' ))

table_file <- file.path(tables_dir, 'Diff_expressed_ERLE.csv')
write_csv(ERLE_to_CTRL.diff_table, table_file )

ERIN_to_CTRL.diff_table <- compounds_table %>% 
  select(-contains('Annotation Source:'), - contains('Results'), -SampleID, -AUC, -Plant, -Fungi) %>% 
  right_join(ERIN_to_CTRL.diff_table, by = 'FeatureID') %>% 
  distinct() %>% 
  filter(!is.nan(log2FC)) %>% 
  mutate(Comment = case_when(log2FC == Inf ~ 'Not present in control',
                             log2FC == -Inf ~ 'Only present in control',
                             log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

ERIN_to_CTRL.diff_table$Comment <- factor(ERIN_to_CTRL.diff_table$Comment, 
                               levels = c('Only present in control', 'Downregulated', 'Upregulated', 'Not present in control' ))

table_file <- file.path(tables_dir, 'Diff_expressed_ERIN.csv')
write_csv(ERIN_to_CTRL.diff_table, table_file )


LEDU_to_CTRL.diff_table <- compounds_table %>% 
  select(-contains('Annotation Source:'), - contains('Results'), -SampleID, -AUC, -Plant, -Fungi) %>% 
  right_join(LEDU_to_CTRL.diff_table, by = 'FeatureID') %>% 
  distinct() %>% 
  filter(!is.nan(log2FC)) %>% 
  mutate(Comment = case_when(log2FC == Inf ~ 'Not present in control',
                             log2FC == -Inf ~ 'Only present in control',
                             log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

LEDU_to_CTRL.diff_table$Comment <- factor(LEDU_to_CTRL.diff_table$Comment, 
                               levels = c('Only present in control', 'Downregulated', 'Upregulated', 'Not present in control' ))

table_file <- file.path(tables_dir, 'Diff_expressed_LEDU.csv')
write_csv(LEDU_to_CTRL.diff_table, table_file )

```

Extract most differentially expressed features

```{r warning=FALSE}

# Extract significant features of each comparison based on adjusted pvalue
sig_features <- c(BOCU_to_CTRL.diff_table$FeatureID[BOCU_to_CTRL.diff_table$pval.adj < 0.05],
                  ERLE_to_CTRL.diff_table$FeatureID[ERLE_to_CTRL.diff_table$pval.adj < 0.05],
                  ERIN_to_CTRL.diff_table$FeatureID[ERIN_to_CTRL.diff_table$pval.adj < 0.05],
                  LEDU_to_CTRL.diff_table$FeatureID[LEDU_to_CTRL.diff_table$pval.adj < 0.05])

sig_features <- unique(sig_features)
```


# 4. Plots of dysregulated features

## 4.1 Number of dysregulated features in each group

```{r}

num_diff_features <- tibble(Comparison = rep(c('BOCU_to_CTRL', 'ERLE_to_CTRL', 'ERIN_to_CTRL', 'LEDU_to_CTRL'), each = 2),
                            Type = rep(c('Name', 'No name'), 4),
                            count = 0)
num_diff_features$Type <- factor(num_diff_features$Type, levels = c('No name', 'Name'))

num_diff_features[1, 3] <- sum(BOCU_to_CTRL.diff_table$pval < 0.05 & !is.na(BOCU_to_CTRL.diff_table$Name), na.rm = TRUE)
num_diff_features[2, 3] <- sum(BOCU_to_CTRL.diff_table$pval < 0.05 & is.na(BOCU_to_CTRL.diff_table$Name), na.rm = TRUE)
num_diff_features[3, 3] <- sum(ERLE_to_CTRL.diff_table$pval < 0.05 & !is.na(ERLE_to_CTRL.diff_table$Name), na.rm = TRUE)
num_diff_features[4, 3] <- sum(ERLE_to_CTRL.diff_table$pval < 0.05 & is.na(ERLE_to_CTRL.diff_table$Name), na.rm = TRUE)
num_diff_features[5, 3] <- sum(ERIN_to_CTRL.diff_table$pval < 0.05 & !is.na(ERIN_to_CTRL.diff_table$Name), na.rm = TRUE)
num_diff_features[6, 3] <- sum(ERIN_to_CTRL.diff_table$pval < 0.05 & is.na(ERIN_to_CTRL.diff_table$Name), na.rm = TRUE)
num_diff_features[5, 3] <- sum(LEDU_to_CTRL.diff_table$pval < 0.05 & !is.na(LEDU_to_CTRL.diff_table$Name), na.rm = TRUE)
num_diff_features[6, 3] <- sum(LEDU_to_CTRL.diff_table$pval < 0.05 & is.na(LEDU_to_CTRL.diff_table$Name), na.rm = TRUE)


#num_diff_plot <- plot_col(num_diff_features, count, Comparison, Type)
#num_diff_plot

#figure_file <- file.path(figures_dir, 'Num_diff_features.png')
#ggsave(figure_file, num_diff_plot, dpi = 300)
```

## 4.2 Volcano plots

```{r warning=FALSE}

lfc.t <- 2
pval.t <- 0.05

BOCU_to_CTRL_volcano <- plot_volcano(BOCU_to_CTRL.diff_table, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'BOCU vs CTRL')
BOCU_to_CTRL_volcano

figure_file <- file.path(figures_dir, 'BOCU_CTRL_volcano.png')
ggsave(figure_file, BOCU_to_CTRL_volcano, dpi=300)


ERLE_to_CTRL_volcano <- plot_volcano(ERLE_to_CTRL.diff_table, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'ERLE vs CTRL')
ERLE_to_CTRL_volcano

figure_file <- file.path(figures_dir, 'ERLE_CTRL_volcano.png')
ggsave(figure_file, ERLE_to_CTRL_volcano, dpi=300)

ERIN_to_CTRL_volcano <- plot_volcano(ERIN_to_CTRL.diff_table, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'ERIN vs CTRL')
ERIN_to_CTRL_volcano

figure_file <- file.path(figures_dir, 'ERIN_CTRL_volcano.png')
ggsave(figure_file, ERIN_to_CTRL_volcano, dpi=300)

LEDU_to_CTRL_volcano <- plot_volcano(LEDU_to_CTRL.diff_table, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'LEDU vs CTRL')
LEDU_to_CTRL_volcano

figure_file <- file.path(figures_dir, 'LEDU_CTRL_volcano.png')
ggsave(figure_file, LEDU_to_CTRL_volcano, dpi=300)
```


```{r}
# Filter diff tables by pval.adj
ERLE.sig <- ERLE_to_CTRL.diff_table %>% 
  filter(pval < 0.05)
ERIN.sig <- ERIN_to_CTRL.diff_table %>% 
  filter(pval < 0.05)
BOCU.sig <- BOCU_to_CTRL.diff_table %>% 
  filter(pval < 0.05)
LEDU.sig <- LEDU_to_CTRL.diff_table %>% 
  filter(pval < 0.05)

# Get list of features at each time point for each of the cases
ERLE.upregulated <- get_vectors(ERLE.sig, 'Comment', 'Upregulated', 'FeatureID')
ERLE.downregulated <- get_vectors(ERLE.sig, 'Comment', 'Downregulated', 'FeatureID')
ERLE.not_in_control <- get_vectors(ERLE.sig, 'Comment', 'Not present in control', 'FeatureID')
ERLE.only_in_control <- get_vectors(ERLE.sig, 'Comment', 'Only present in control', 'FeatureID')
ERIN.upregulated <- get_vectors(ERIN.sig, 'Comment', 'Upregulated', 'FeatureID')
ERIN.downregulated <- get_vectors(ERIN.sig, 'Comment', 'Downregulated', 'FeatureID')
ERIN.not_in_control <- get_vectors(ERIN.sig, 'Comment', 'Not present in control', 'FeatureID')
ERIN.only_in_control <- get_vectors(ERIN.sig, 'Comment', 'Only present in control', 'FeatureID')
BOCU.upregulated <- get_vectors(BOCU.sig, 'Comment', 'Upregulated', 'FeatureID')
BOCU.downregulated <- get_vectors(BOCU.sig, 'Comment', 'Downregulated', 'FeatureID')
BOCU.not_in_control <- get_vectors(BOCU.sig, 'Comment', 'Not present in control', 'FeatureID')
BOCU.only_in_control <- get_vectors(BOCU.sig, 'Comment', 'Only present in control', 'FeatureID')
LEDU.upregulated <- get_vectors(LEDU.sig, 'Comment', 'Upregulated', 'FeatureID')
LEDU.downregulated <- get_vectors(LEDU.sig, 'Comment', 'Downregulated', 'FeatureID')
LEDU.not_in_control <- get_vectors(LEDU.sig, 'Comment', 'Not present in control', 'FeatureID')
LEDU.only_in_control <- get_vectors(LEDU.sig, 'Comment', 'Only present in control', 'FeatureID')
```

```{r}

ERLE.count<- ERLE.sig%>% 
    select(FeatureID, Class, Comment, pval, pval.adj)%>%
    count(Class, Comment)%>%
    filter(Comment != "Only present in control")
  
class.plot<- ggplot(ERLE.count)+
geom_bar(aes(y=Class, x=n,fill= Comment), stat="identity", alpha=0.7)+
  #scale_fill_manual(values = c("lightskyblue","firebrick2"))+
  theme_classic()+xlab("Number of Factors")+ylab("Class")+
  labs(title="Class ranks of Isolate CTRL")
class.plot

figure_file <- file.path(figures_dir, 'Class_rank_ERLE.png')
ggsave(figure_file, class.plot, width=15, height=10, unit='cm')


BOCU.count<- BOCU.sig%>% 
    select(FeatureID, Class, Comment, pval, pval.adj)%>%
    count(Class, Comment)%>%
    filter(Comment != "Only present in control")
  
class.plot<- ggplot(BOCU.count)+
geom_bar(aes(y=Class, x=n,fill= Comment), stat="identity", alpha=0.7)+
  #scale_fill_manual(values = c("lightskyblue","firebrick2"))+
  theme_classic()+xlab("Number of Factors")+ylab("Class")+
  labs(title="Class ranks of Isolate CTRL")
class.plot

figure_file <- file.path(figures_dir, 'Class_rank_BOCU.png')
ggsave(figure_file, class.plot, width=15, height=10, unit='cm')

ERIN.count<- ERIN.sig%>% 
    select(FeatureID, Class, Comment, pval, pval.adj)%>%
    count(Class, Comment)%>%
    filter(Comment != "Only present in control")
  
class.plot<- ggplot(ERIN.count)+
geom_bar(aes(y=Class, x=n,fill= Comment), stat="identity", alpha=0.7)+
  #scale_fill_manual(values = c("lightskyblue","firebrick2"))+
  theme_classic()+xlab("Number of Factors")+ylab("Class")+
  labs(title="Class ranks of Isolate CTRL")
class.plot

figure_file <- file.path(figures_dir, 'Class_rank_ERIN.png')
ggsave(figure_file, class.plot, width=15, height=10, unit='cm')

LEDU.count<- LEDU.sig%>% 
    select(FeatureID, Class, Comment, pval, pval.adj)%>%
    count(Class, Comment)%>%
    filter(Comment != "Only present in control")
  
class.plot<- ggplot(LEDU.count)+
geom_bar(aes(y=Class, x=n,fill= Comment), stat="identity", alpha=0.7)+
  #scale_fill_manual(values = c("lightskyblue","firebrick2"))+
  theme_classic()+xlab("Number of Factors")+ylab("Class")+
  labs(title="Class ranks of Isolate CTRL")
class.plot

figure_file <- file.path(figures_dir, 'Class_rank_LEDU.png')
ggsave(figure_file, class.plot, width=15, height=10, unit='cm')
```


```{r}
BOCU.fold<- BOCU.sig %>% 
  select(FeatureID, name4plot, KEGGPathways, Formula, MW, Comment, log2FC, pval)%>% 
  filter(log2FC> 2| log2FC<(-2))

table_file <- file.path(tables_dir, 'Sig.FoldChange_BOCU.csv')
write_csv( BOCU.fold, table_file)
  

LEDU.fold<- LEDU.sig%>% 
 select(FeatureID, name4plot, KEGGPathways, Formula, MW, Comment, log2FC, pval)%>%
  separate_rows(KEGGPathways, sep = ';') %>% 
   filter(log2FC> 2| log2FC<(-2))

table_file <- file.path(tables_dir, 'Sig.FoldChange_LEDU.csv')
write_csv( LEDU.fold, table_file)

ERLE.fold<- ERLE.sig%>% 
 select(FeatureID, name4plot, KEGGPathways, Formula, MW, Comment, log2FC, pval)%>%
  separate_rows(KEGGPathways, sep = ';') %>% 
  filter(log2FC> 2| log2FC<(-2))
 
table_file <- file.path(tables_dir, 'Sig.FoldChange_ERIN.csv')
write_csv( ERLE.fold, table_file)

 
ERIN.fold<- ERIN.sig%>% 
  select(FeatureID, name4plot, KEGGPathways, Formula, MW, Comment, log2FC, pval)%>% 
  separate_rows(KEGGPathways, sep = ';') %>% 
 filter(log2FC> 2| log2FC<(-2))
 
table_file <- file.path(tables_dir, 'Sig.FoldChange_ERLE.csv')
write_csv( ERIN.fold, table_file)
```


# 4.38 Upregulated Secondary Metabolites
```{r}

BOCU.sig$KEGGPathways<-as.factor(BOCU.sig$"KEGG Pathways")  
BOCU.reg<- BOCU.sig %>% 
  #select(FeatureID, KEGGPathways, Comment) %>% 
  separate_rows(KEGGPathways, sep = ';') %>% 
  filter(KEGGPathways== 'Biosynthesis of secondary metabolites')%>% 
 filter(Comment == "Upregulated")

table_file <- file.path(tables_dir, 'Upreg.SecondaryMetabolites_BOCU.csv')
write_csv( BOCU.reg, table_file)
  
LEDU.sig$KEGGPathways<-as.factor(LEDU.sig$"KEGG Pathways")  
LEDU.reg<- LEDU.sig%>% 
  #select(FeatureID, KEGGPathways, Comment) %>% 
  separate_rows(KEGGPathways, sep = ';') %>% 
  filter(KEGGPathways== 'Biosynthesis of secondary metabolites')%>% 
 filter(Comment == "Upregulated")
table_file <- file.path(tables_dir, 'Upreg.SecondaryMetabolites_LEDU.csv')
write_csv( LEDU.reg, table_file)

ERLE.sig$KEGGPathways<-as.factor(ERLE.sig$"KEGG Pathways")  
ERLE.reg<- ERLE.sig%>% 
  #select(FeatureID, KEGGPathways, Comment) %>% 
  separate_rows(KEGGPathways, sep = ';') %>% 
  filter(KEGGPathways== 'Biosynthesis of secondary metabolites')%>% 
 filter(Comment == "Upregulated")
table_file <- file.path(tables_dir, 'Upreg.SecondaryMetabolites_ERIN.csv')
write_csv( ERLE.reg, table_file)

ERIN.sig$KEGGPathways<-as.factor(ERIN.sig$"KEGG Pathways")  
ERIN.reg<- ERIN.sig%>% 
  #select(FeatureID, KEGGPathways, Comment) %>% 
  separate_rows(KEGGPathways, sep = ';') %>% 
  filter(KEGGPathways== 'Biosynthesis of secondary metabolites')%>% 
 filter(Comment == "Upregulated")
table_file <- file.path(tables_dir, 'Upreg.SecondaryMetabolites_ERIN.csv')
write_csv( ERIN.reg, table_file)
```

## 4.3 KEGG Up and Down Regulated 
```{r}

ERLE.sig$KEGGPathways<-as.factor(ERLE.sig$"KEGG Pathways")  
ERLE.count<- ERLE.sig%>% 
  select(FeatureID, KEGGPathways, Comment) %>% 
  separate_rows(KEGGPathways, sep = ';') %>% 
  group_by(KEGGPathways, Comment ) %>% 
  count(KEGGPathways) %>%
   drop_na(KEGGPathways)%>% filter(Comment != "Only present in control")
  
class.plot<- ggplot(ERLE.count)+
geom_bar(aes(y=KEGGPathways, x=n,fill= Comment), stat="identity", alpha=0.7)+
  #scale_fill_manual(values = c("lightskyblue","firebrick2"))+
  theme_classic()+xlab("Number of Factors")+ylab("KEGG Pathways")+
  labs(title="Up and down regulated KEGG pathways of Isolate CTRL")
class.plot

figure_file <- file.path(figures_dir, 'KEGGrank_ERLE.png')
ggsave(figure_file, class.plot, width=15, height=10, unit='cm')


BOCU.sig$KEGGPathways<-as.factor(BOCU.sig$"KEGG Pathways")  
BOCU.count<- BOCU.sig%>% 
  select(FeatureID, KEGGPathways, Comment) %>% 
  separate_rows(KEGGPathways, sep = ';') %>% 
  group_by(KEGGPathways, Comment ) %>% 
  count(KEGGPathways) %>%
   drop_na(KEGGPathways)%>% filter(Comment != "Only present in control")
  
class.plot<- ggplot(BOCU.count)+
geom_bar(aes(y=KEGGPathways, x=n,fill= Comment), stat="identity", alpha=0.7)+
  #scale_fill_manual(values = c("lightskyblue","firebrick2"))+
  theme_classic()+xlab("Number of Factors")+ylab("KEGG Pathways")+
  labs(title="Up and down regulated KEGG pathways of Isolate CTRL")
class.plot

figure_file <- file.path(figures_dir, 'KEGGrank_BOCU.png')
ggsave(figure_file, class.plot, width=15, height=10, unit='cm')


ERIN.sig$KEGGPathways<-as.factor(ERIN.sig$"KEGG Pathways")  
ERIN.count<- ERIN.sig%>% 
  select(FeatureID, KEGGPathways, Comment) %>% 
  separate_rows(KEGGPathways, sep = ';') %>% 
  group_by(KEGGPathways, Comment ) %>% 
  count(KEGGPathways) %>%
   drop_na(KEGGPathways)%>% filter(Comment != "Only present in control")
  
class.plot<- ggplot(ERIN.count)+
geom_bar(aes(y=KEGGPathways, x=n,fill= Comment), stat="identity", alpha=0.7)+
  #scale_fill_manual(values = c("lightskyblue","firebrick2"))+
  theme_classic()+xlab("Number of Factors")+ylab("KEGG Pathways")+
  labs(title="Up and down regulated KEGG pathways of Isolate CTRL")
class.plot

figure_file <- file.path(figures_dir, 'KEGGrank_ERIN.png')
ggsave(figure_file, class.plot, width=15, height=10, unit='cm')
```



## 4.38 Upregulated Secondary Metabolites
```{r}

ERLE.sig$KEGGPathways<-as.factor(ERLE.sig$"KEGG Pathways")  
ERLE.reg<- ERLE.sig %>% 
  #select(FeatureID, KEGGPathways, Comment) %>% 
  separate_rows(KEGGPathways, sep = ';') %>% 
  filter(KEGGPathways== 'Biosynthesis of secondary metabolites')%>% 
 filter(Comment == "Upregulated")

table_file <- file.path(tables_dir, 'Upreg.SecondaryMetabolites_ERIN.csv')
write_csv( ERLE.reg, table_file)
  
BOCU.sig$KEGGPathways<-as.factor(BOCU.sig$"KEGG Pathways")  
BOCU.reg<- BOCU.sig%>% 
  #select(FeatureID, KEGGPathways, Comment) %>% 
  separate_rows(KEGGPathways, sep = ';') %>% 
  filter(KEGGPathways== 'Biosynthesis of secondary metabolites')%>% 
 filter(Comment == "Upregulated")

ERIN.sig$KEGGPathways<-as.factor(ERIN.sig$"KEGG Pathways")  
ERIN.reg<- ERIN.sig%>% 
  #select(FeatureID, KEGGPathways, Comment) %>% 
  separate_rows(KEGGPathways, sep = ';') %>% 
  filter(KEGGPathways== 'Biosynthesis of secondary metabolites')%>% 
 filter(Comment == "Upregulated")
```







# 5. Hierarchical clustering analysis using heatmaps

Multiple heatmaps will be plotted, based on the differential expression and if the compounds were or not assigned and structure

### Normalized Area under the curve (AUC) heatmap for all detected features

```{r warning=FALSE}

# Initialize graphical device
#dev.off()

# Set sampleID as row.names to annotate heatmap

col_annot <- column_to_rownames(metadata, var = 'SampleID') 
mapcolor <- colorRampPalette(brewer.pal(11, 'RdYlBu'))(100)[100:1]

figure_file <- file.path(figures_dir, 'All_features_heatmap.pdf')

annot_colors <- list(
  Time = c(CTRL ='#0073C2FF', ERIN = '#EFC000FF', BOCU = '#868686FF', ERLE = '#CD534CFF'),
  Material = c(Litter = 'green4', Litter_Peat = 'chocolate4'),
  Origin = c(LC_MS2 = '#0086cb', NMR = '#d23936'))

pdf(figure_file)
pheatmap(norm.matrix,
         clustering_distance_rows = 'correlation',
         clustering_distance_cols = 'correlation',
         scale = 'row',
         annotation_col = col_annot,
         annotation_colors = annot_colors,
         color = mapcolor,
         show_rownames = FALSE,
         cutree_cols = 5,
         main = 'All features (scaled AUC)'
)
dev.off()
```

```{r}
ERLE.sig$Plant<- 'ERLE'
ERIN.sig$Plant<- 'ERIN'
BOCU.sig$Plant<- 'BOCU'
LEDU.sig$Plant<- 'LEDU'

sig.data.all<- rbind(ERLE.sig, ERIN.sig, BOCU.sig, LEDU.sig)
sig.data.all$Class <- gsub( "Lignin", "Phenol", sig.data.all$Class)
sig.data.all$Class <- gsub("Tannin","Other",  sig.data.all$Class)


table_file <- file.path(tables_dir, 'sig.data.all.csv')
write.csv(sig.data.all, table_file, row.names = TRUE)

sig.data.duplicate<- sig.data.all%>%
  filter(duplicated(FeatureID), keep.all=TRUE)
table_file <- file.path(tables_dir, 'sig.data.duplicates.csv')
write.csv(sig.data.duplicate, table_file, row.names = TRUE)

sig.data.all<- sig.data.all %>%
  select(FeatureID, Class, Plant, Comment)%>%
  count(Class, Plant, Comment)

stack.class<-ggplot(sig.data.all, aes(fill=Class, y=n, x=Plant)) + 
    geom_bar(position="fill", stat="identity")+
  facet_wrap(~Comment)+
  scale_fill_brewer(palette = 'Set2')+
  theme_classic()+
  ylab('Percent abundance of compounds')+
  xlab('Fungal isolates')+
  labs(title= "Differentially expressed compounds by compound class of isolate 1154",
       caption="Differetally expressed compounds pval.adj< 0.05 ")
stack.class
figure_file <- file.path(figures_dir, 'Class.Percent.abund.png')
ggsave(figure_file, stack.class, dpi=400)

stack.class2<-ggplot(sig.data.all, aes(fill=Class, y=n, x=Plant)) + 
    geom_bar(position="stack", stat="identity")+
  facet_wrap(~Comment)+
  scale_fill_brewer(palette = 'Set2')+
  theme_classic()+
  ylab('Number of compounds')+
  xlab('Fungal isolates')+
  labs(title= "Differentially expressed compounds by compound class of isolate 1154",
       caption="Differetally expressed compounds pval.adj< 0.05 ")
stack.class2
figure_file <- file.path(figures_dir, 'Class.Relative.abund.png')
ggsave(figure_file, stack.class2, dpi=400)
```

[Scaled Area under the curve (AUC) heatmap for all detected features](`r toString(figure_file)`)

### Normalized Area under the curve (AUC) heatmap for only features that were identified

```{r}
# Extract all features that have names
named_compounds <- compounds_table %>% 
  select(FeatureID, Name, name4plot) %>% 
  filter(!is.na(Name)) %>% 
  distinct()

# Create a matrix of only identified features

named_compounds.matrix <- norm.matrix[rownames(norm.matrix) %in% named_compounds$FeatureID,]
row.names(named_compounds.matrix) <- compounds_table$name4plot[match(row.names(named_compounds.matrix), compounds_table$FeatureID)]

```

```{r warning=FALSE}

figure_file <- file.path(figures_dir, 'All_identified_features_heatmap.pdf')

pdf(figure_file, width = 15, height = 10)
pheatmap(named_compounds.matrix,
         clustering_distance_rows = 'correlation',
         clustering_distance_cols = 'correlation',
         scale = 'row',
         annotation_col = col_annot,
         annotation_colors = annot_colors,
         color = mapcolor,
         cutree_cols = 4,
         cutree_rows = 5,
         fontsize_row = 6,
         main = 'Identified features (scaled AUC)'
)
dev.off()
```

[Scaled Area under the curve (AUC) heatmap for only features that were identified](`r toString(figure_file)`)

### All dysregulated features

```{r}
# Create an AUC matrix that contains only features that are differentially expressed

sig_features.matrix <- norm.matrix[row.names(norm.matrix) %in% sig_features, ]

row.names(sig_features.matrix) <- compounds_table$name4plot[match(row.names(sig_features.matrix), compounds_table$FeatureID)]
```



[Scaled Area under the curve (AUC) heatmap for all dysregulated features](`r toString(figure_file)`)

### Dysregulated features that were identified

```{r }
named_sig_features.matrix <- sig_features.matrix[rownames(sig_features.matrix) %in% named_compounds$Name, ]

figure_file <- file.path(figures_dir, 'Dysreg_identified_features.pdf')

pdf(figure_file, width = 10)
pheatmap(named_sig_features.matrix,
         clustering_distance_rows = 'correlation',
         clustering_distance_cols = 'correlation',
         scale = 'row',
         annotation_col = col_annot,
         annotation_colors = annot_colors,
         color = mapcolor,
         cutree_cols = 4,
         fontsize_row = 8,
         main = 'Identified dysregulated features (scaled AUC)'
)
dev.off()
```




## 4.3 Van Krevelen Diagrams of dysregulated features

**Dysregulated features at BOCU**

```{r}

BOCU_vank <- plot_vank(BOCU.sig, Class, Comment)

BOCU_vank

figure_file <- file.path(figures_dir, 'DE_vank_BOCU.png')
ggsave(figure_file, BOCU_vank, dpi=300, unit='cm')
```

**Dysregulated features at CTRL**

```{r}
ERLE_vank <- plot_vank(ERLE.sig, Class, Comment)
ERLE_vank

figure_file <- file.path(figures_dir, 'DE_vank_ERLE.png')
ggsave(figure_file, ERLE_vank,dpi=300, unit='cm')
```

**Dysregulated features at ERIN**

```{r}
ERIN_vank <- plot_vank(ERIN.sig, Class, Comment)
ERIN_vank

figure_file <- file.path(figures_dir, 'DE_vank_ERIN.png')
ggsave(figure_file, ERIN_vank,dpi=300, unit='cm')
```
```{r}
LEDU_vank <- plot_vank(LEDU.sig, Class, Comment)
LEDU_vank

figure_file <- file.path(figures_dir, 'DE_vank_LEDU.png')
ggsave(figure_file, LEDU_vank,dpi=300, unit='cm')
```