---
title: "DifferentialAnalysis"
output: html_notebook
---

---
title: "RP Differential Analysis"
output: html_notebook
---

# 1. Importing Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggrepel)
library(ggsci)
library(pheatmap)
library(RColorBrewer)
library(readxl)
library(UpSetR)
source('functions_cdis_diff.R')
```

Change to what data you are using
```{r}
data= 'RP'
```

# 2. Import data


Because differential analysis includes the calculations of means, the data to be used in this section is the normalized untransformed data

```{r set_path, message=FALSE, warning=FALSE}
# set path variables
project_dir <- getwd()
project_name <- 'RP_updted_parameters'
figures_dir <- file.path(project_dir, paste0(project_name, '_output_figures'))
tables_dir <- file.path(project_dir,  paste0(project_name, '_output_tables'))
norm_auc_table_file <- file.path(tables_dir, 'normalized_untransformed_auc_table.csv')

# If the flag was setted before, nothing needs to be changed here, the correct file will be automatically used

  compounds_table_file <- file.path(tables_dir, 'compounds_table.csv')


# Load auc_table

norm.matrix <- read_csv(norm_auc_table_file)
norm.matrix <- column_to_rownames(norm.matrix, var = '...1')

# Load compounds table and add a column that has names for identified compounds and FeatureID for the rest (for plotting puRPouses)

compounds_table <- read_csv(compounds_table_file)

# Import metadata and fix names
metadata_file <- file.path(tables_dir, 'fixed_metadata.csv')
metadata <- read_csv(metadata_file)



```



# 3. Calculate ratios and log2fold-change

## 3.1 Get samples per each treatment

```{r}

# Get the average AUC per each of the treatments
## Get samples per treatment

F1154.samples <- get_samples(metadata, Treatment = 'Fungi', value = '1154')
F1177.samples <- get_samples(metadata, Treatment = 'Fungi', value = '1177')
F1246.samples <- get_samples(metadata, Treatment = 'Fungi', value = '1246')
CTRL.samples <- get_samples(metadata, Treatment = 'Fungi', value = 'CTRL')

```

## 3.2 Get table of differentially expressed features per each comparison

The following function will calculate ratio, log2FC, p values and adjusted pvalues. If no replicates are available for EACH treatment please use the get_diff_table_no_pval() function

```{r}

# The following function will calculate ratio, log2FC, p values and adjusted pvalues. If no replicates are available for EACH treatment
# please use the get_diff_table_no_pval() function

F1154_to_CTRL.diff_table <- get_diff_table(norm.matrix, treatment.sample_list = F1154.samples, control.sample_list = CTRL.samples)
F1177_to_CTRL.diff_table <- get_diff_table(norm.matrix, treatment.sample_list = F1177.samples, control.sample_list = CTRL.samples)
F1246_to_CTRL.diff_table <- get_diff_table(norm.matrix, treatment.sample_list = F1246.samples, control.sample_list = CTRL.samples)
```

Merge differentially expressed features with the annotation

```{r}

# Create dataframes for the up and downregulated metabolites at each time point and merge them with the compound information

F1154_to_CTRL.diff_table <- compounds_table %>% 
  select( - contains('Results'), -SampleID, -AUC,  -Fungi, - Plant) %>% 
  right_join(F1154_to_CTRL.diff_table, by = 'FeatureID') %>% 
  distinct() %>% 
  filter(!is.nan(log2FC)) %>% 
  mutate(Comment = case_when(log2FC == Inf ~ 'Not present in control',
                             log2FC == -Inf ~ 'Only present in control',
                             log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

F1154_to_CTRL.diff_table$Comment <- factor(F1154_to_CTRL.diff_table$Comment, 
                               levels = c('Only present in control', 'Downregulated', 'Upregulated', 'Not present in control' ))

table_file <- file.path(tables_dir, 'Diff_expressed_F1154.csv')
write_csv(F1154_to_CTRL.diff_table, table_file )

F1177_to_CTRL.diff_table<- compounds_table %>% 
  select(-contains('Annotation Source:'), - contains('Results'), -SampleID, -AUC, -Plant, - Fungi) %>%
  right_join(F1177_to_CTRL.diff_table, by = 'FeatureID') %>% 
  distinct() %>% 
  filter(!is.nan(log2FC)) %>% 
  mutate(Comment = case_when(log2FC == Inf ~ 'Not present in control',
                             log2FC == -Inf ~ 'Only present in control',
                             log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

F1177_to_CTRL.diff_table$Comment <- factor(F1177_to_CTRL.diff_table$Comment, 
                               levels = c('Only present in control', 'Downregulated', 'Upregulated', 'Not present in control' ))

table_file <- file.path(tables_dir, 'Diff_expressed_F1177.csv')
write_csv(F1177_to_CTRL.diff_table, table_file )

F1246_to_CTRL.diff_table <- compounds_table %>% 
  select(-contains('Annotation Source:'), - contains('Results'), -SampleID, -AUC, -Plant, -Fungi) %>% 
  right_join(F1246_to_CTRL.diff_table, by = 'FeatureID') %>% 
  distinct() %>% 
  filter(!is.nan(log2FC)) %>% 
  mutate(Comment = case_when(log2FC == Inf ~ 'Not present in control',
                             log2FC == -Inf ~ 'Only present in control',
                             log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

F1246_to_CTRL.diff_table$Comment <- factor(F1246_to_CTRL.diff_table$Comment, 
                               levels = c('Only present in control', 'Downregulated', 'Upregulated', 'Not present in control' ))

table_file <- file.path(tables_dir, 'Diff_expressed_F1246.csv')
write_csv(F1246_to_CTRL.diff_table, table_file )

```

Extract most differentially expressed features

```{r warning=FALSE}

# Extract significant features of each comparison based on adjusted pvalue
sig_features <- c(F1154_to_CTRL.diff_table$FeatureID[F1154_to_CTRL.diff_table$pval.adj < 0.05],
                  F1177_to_CTRL.diff_table$FeatureID[F1177_to_CTRL.diff_table$pval.adj < 0.05],
                  F1246_to_CTRL.diff_table$FeatureID[F1246_to_CTRL.diff_table$pval.adj < 0.05])

sig_features <- unique(sig_features)

```


# 4. Plots of dysregulated features

## 4.1 Number of dysregulated features in each group

```{r}
# removed pval adj to regular pval
num_diff_features <- tibble(Comparison = rep(c('F1154_to_CTRL', 'F1177_to_CTRL', 'F1246_to_CTRL'), each = 2),
                            Type = rep(c('Name', 'No name'), 3),
                            count = 0)
num_diff_features$Type <- factor(num_diff_features$Type, levels = c('No name', 'Name'))

num_diff_features[1, 3] <- sum(F1154_to_CTRL.diff_table$pval.adj < 0.05 & !is.na(F1154_to_CTRL.diff_table$Name), na.rm = TRUE)
num_diff_features[2, 3] <- sum(F1154_to_CTRL.diff_table$pval.adj < 0.05 & is.na(F1154_to_CTRL.diff_table$Name), na.rm = TRUE)
num_diff_features[3, 3] <- sum(F1177_to_CTRL.diff_table$pval.adj < 0.05 & !is.na(F1177_to_CTRL.diff_table$Name), na.rm = TRUE)
num_diff_features[4, 3] <- sum(F1177_to_CTRL.diff_table$pval.adj < 0.05 & is.na(F1177_to_CTRL.diff_table$Name), na.rm = TRUE)
num_diff_features[5, 3] <- sum(F1246_to_CTRL.diff_table$pval.adj < 0.05 & !is.na(F1246_to_CTRL.diff_table$Name), na.rm = TRUE)
num_diff_features[6, 3] <- sum(F1246_to_CTRL.diff_table$pval.adj < 0.05 & is.na(F1246_to_CTRL.diff_table$Name), na.rm = TRUE)


num_diff_plot <- ggplot(num_diff_features)+
  geom_bar(aes(y=count, x=Comparison,fill= Type), stat="identity", alpha=0.7)+
  theme_classic()+xlab("Comparison")+ylab("Count")
  
num_diff_plot
# shows snumber of significantly regulated compounds
figure_file <- file.path(figures_dir, 'Num_diff_features.png')
ggsave(figure_file, num_diff_plot, dpi = 300)
```

```{r warning=FALSE}

# Extract significant features of each comparison based on adjusted pvalue
sig_features <- c(F1154_to_CTRL.diff_table$FeatureID[F1154_to_CTRL.diff_table$pval.adj < 0.05],
                  F1177_to_CTRL.diff_table$FeatureID[F1177_to_CTRL.diff_table$pval.adj < 0.05],
                  F1246_to_CTRL.diff_table$FeatureID[F1246_to_CTRL.diff_table$pval.adj < 0.05])

sig_features <- unique(sig_features)
```


# 4. Plots of dysregulated features

## 4.1 Number of dysregulated features in each group

```{r}

num_diff_features <- tibble(Comparison = rep(c('F1154_to_CTRL', 'F1177_to_CTRL', 'F1246_to_CTRL'), each = 2),
                            Type = rep(c('Name', 'No name'), 3),
                            count = 0)
num_diff_features$Type <- factor(num_diff_features$Type, levels = c('No name', 'Name'))

num_diff_features[1, 3] <- sum(F1154_to_CTRL.diff_table$pval.adj < 0.05 & !is.na(F1154_to_CTRL.diff_table$Name), na.rm = TRUE)
num_diff_features[2, 3] <- sum(F1154_to_CTRL.diff_table$pval.adj < 0.05 & is.na(F1154_to_CTRL.diff_table$Name), na.rm = TRUE)
num_diff_features[3, 3] <- sum(F1177_to_CTRL.diff_table$pval.adj < 0.05 & !is.na(F1177_to_CTRL.diff_table$Name), na.rm = TRUE)
num_diff_features[4, 3] <- sum(F1177_to_CTRL.diff_table$pval.adj < 0.05 & is.na(F1177_to_CTRL.diff_table$Name), na.rm = TRUE)
num_diff_features[5, 3] <- sum(F1246_to_CTRL.diff_table$pval.adj < 0.05 & !is.na(F1246_to_CTRL.diff_table$Name), na.rm = TRUE)
num_diff_features[6, 3] <- sum(F1246_to_CTRL.diff_table$pval.adj < 0.05 & is.na(F1246_to_CTRL.diff_table$Name), na.rm = TRUE)


num_diff_plot <- ggplot(num_diff_features)+
  geom_bar(aes(y=count, x=Comparison,fill= Type), stat="identity", alpha=0.7)+
  theme_classic()+xlab("Comparison")+ylab("Count")
  
num_diff_plot

figure_file <- file.path(figures_dir, 'Num_diff_features.png')
ggsave(figure_file, num_diff_plot, dpi = 300)
```



## 4.2 Volcano plots

```{r warning=FALSE}

lfc.t <- 1
pval.t <- 0.05

F1154_to_CTRL_volcano <- plot_volcano(F1154_to_CTRL.diff_table, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'Isolate 1154 vs Media')
F1154_to_CTRL_volcano

figure_file <- file.path(figures_dir, 'F1154_CTRL_volcano.png')
ggsave(figure_file, F1154_to_CTRL_volcano, width=20, height=20, unit='cm')


F1177_to_CTRL_volcano <- plot_volcano(F1177_to_CTRL.diff_table, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'Isolate 1177 vs Media')
F1177_to_CTRL_volcano

figure_file <- file.path(figures_dir, 'F1177_CTRL_volcano.png')
ggsave(figure_file, F1177_to_CTRL_volcano, width=20, height=20, unit='cm')

F1246_to_CTRL_volcano <- plot_volcano(F1246_to_CTRL.diff_table, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'Isolate 1246 vs Media')
F1246_to_CTRL_volcano

figure_file <- file.path(figures_dir, 'F1246_CTRL_volcano.png')
ggsave(figure_file, F1246_to_CTRL_volcano, width=20, height=20, unit='cm')
```


```{r}
# Filter diff tables by pval.adj
F1246.sig <- F1246_to_CTRL.diff_table %>% 
  filter(pval.adj < 0.05)
F1154.sig <- F1154_to_CTRL.diff_table %>% 
  filter(pval.adj < 0.05)
F1177.sig <- F1177_to_CTRL.diff_table %>% 
  filter(pval.adj < 0.05)
# Get list of features at each time point for each of the cases
F1246.upregulated <- get_vectors(F1246.sig, 'Comment', 'Upregulated', 'FeatureID')
F1246.downregulated <- get_vectors(F1246.sig, 'Comment', 'Downregulated', 'FeatureID')
F1246.not_in_control <- get_vectors(F1246.sig, 'Comment', 'Not present in control', 'FeatureID')
F1246.only_in_control <- get_vectors(F1246.sig, 'Comment', 'Only present in control', 'FeatureID')
F1154.upregulated <- get_vectors(F1154.sig, 'Comment', 'Upregulated', 'FeatureID')
F1154.downregulated <- get_vectors(F1154.sig, 'Comment', 'Downregulated', 'FeatureID')
F1154.not_in_control <- get_vectors(F1154.sig, 'Comment', 'Not present in control', 'FeatureID')
F1154.only_in_control <- get_vectors(F1154.sig, 'Comment', 'Only present in control', 'FeatureID')
F1177.upregulated <- get_vectors(F1177.sig, 'Comment', 'Upregulated', 'FeatureID')
F1177.downregulated <- get_vectors(F1177.sig, 'Comment', 'Downregulated', 'FeatureID')
F1177.not_in_control <- get_vectors(F1177.sig, 'Comment', 'Not present in control', 'FeatureID')
F1177.only_in_control <- get_vectors(F1177.sig, 'Comment', 'Only present in control', 'FeatureID')
```



```{r}

F1246.count<- F1246.sig%>% 
    select(FeatureID, Class, Comment, pval.adj)%>%
    count(Class, Comment)

F1246.count <- within(F1246.count, n[Comment=="Downregulated"] <- (n*(-1)))

  
class.plot<- ggplot(F1246.count)+
geom_bar(aes(y=Class, x=n,fill= Comment), stat="identity", alpha=0.7)+
  #scale_fill_manual(values = c("lightskyblue","firebrick2"))+
  theme_classic()+xlab("Number of Factors")+ylab("Class")+
  labs(title="Class ranks of Isolate 1246")
class.plot

figure_file <- file.path(figures_dir, 'Class_rank_F1246.png')
ggsave(figure_file, class.plot, width=15, height=10, unit='cm')


F1177.count<- F1177.sig%>% 
    select(FeatureID, Class, Comment, pval.adj)%>%
    count(Class, Comment)
F1177.count <- within(F1177.count, n[Comment=="Downregulated"] <- (n*(-1)))
    #filter(Comment != "Only present in control")
  
class.plot<- ggplot(F1177.count)+
geom_bar(aes(y=Class, x=n,fill= Comment), stat="identity", alpha=0.7)+
  #scale_fill_manual(values = c("lightskyblue","firebrick2"))+
  theme_classic()+xlab("Number of Factors")+ylab("Class")+
  labs(title="Class ranks of Isolate 1177")
class.plot

figure_file <- file.path(figures_dir, 'Class_rank_F1177.png')
ggsave(figure_file, class.plot, width=15, height=10, unit='cm')

F1154.count<- F1154.sig%>% 
    select(FeatureID, Class, Comment, pval.adj, pval)%>%
    count(Class, Comment)
F1154.count <- within(F1154.count, n[Comment=="Downregulated"] <- (n*(-1)))
    #filter(Comment != "Only present in control")
  
class.plot<- ggplot(F1154.count)+
geom_bar(aes(y=Class, x=n,fill= Comment), stat="identity", alpha=0.7)+
  #scale_fill_manual(values = c("lightskyblue","firebrick2"))+
  theme_classic()+xlab("Number of Factors")+ylab("Class")+
  labs(title="Class ranks of Isolate 1154")
class.plot

figure_file <- file.path(figures_dir, 'Class_rank_F1154.png')
ggsave(figure_file, class.plot, width=15, height=10, unit='cm')


```






# 5. Hierarchical clustering analysis using heatmaps

Multiple heatmaps will be plotted, based on the differential expression and if the compounds were or not assigned and structure

### Normalized Area under the curve (AUC) heatmap for all detected features

```{r warning=FALSE}

# Initialize graphical device
#dev.off()

# Set sampleID as row.names to annotate heatmap

col_annot <- column_to_rownames(metadata, var = 'SampleID') 
mapcolor <- colorRampPalette(brewer.pal(11, 'RdYlBu'))(100)[100:1]

figure_file <- file.path(figures_dir, 'All_features_heatmap.pdf')

annot_colors <- list(
  Time = c(CTRL ='#0073C2FF', F1154 = '#EFC000FF', F1177 = '#868686FF', F1246 = '#CD534CFF'),
  Material = c(Litter = 'green4', Litter_Peat = 'chocolate4'),
  Origin = c(LC_MS2 = '#0086cb', NMR = '#d23936'))

pdf(figure_file)
pheatmap(norm.matrix,
         clustering_distance_rows = 'correlation',
         clustering_distance_cols = 'correlation',
         scale = 'row',
         annotation_col = col_annot,
         annotation_colors = annot_colors,
         color = mapcolor,
         show_rownames = FALSE,
         cutree_cols = 5,
         main = 'All features (scaled AUC)'
)
#dev.off()
```

[Scaled Area under the curve (AUC) heatmap for all detected features](`r toString(figure_file)`)

### Normalized Area under the curve (AUC) heatmap for only features that were identified

```{r}
# Extract all features that have names
named_compounds <- compounds_table %>% 
  select(FeatureID, Name, name4plot) %>% 
  filter(!is.na(Name)) %>% 
  distinct()

# Create a matrix of only identified features

named_compounds.matrix <- norm.matrix[rownames(norm.matrix) %in% named_compounds$FeatureID,]
row.names(named_compounds.matrix) <- compounds_table$name4plot[match(row.names(named_compounds.matrix), compounds_table$FeatureID)]

```


```{r}

F1154.sig$Fungi<- '1154'
F1177.sig$Fungi<- '1177'
F1246.sig$Fungi<- '1246'


sig.data.all<- rbind(F1154.sig, F1177.sig, F1246.sig)
#compounds_table$Class <- gsub( "Lignin", "Phenol", compounds_table$Class)
sig.data.all$Class <- gsub("Tannin","Other",  sig.data.all$Class)

sig.data.all<- sig.data.all %>%
  mutate(class = case_when(
    endsWith(Class, "Lignin") ~ "Phenol",
    endsWith(Class, "Tannin") ~ "Highly oxygenated compounds"
  ))%>%
  select(FeatureID, Class, Fungi, Comment)%>%
  count(Class, Fungi, Comment)
  


stack.class<-ggplot(sig.data.all, aes(fill=Class, y=n, x=Fungi)) + 
    facet_wrap(~Comment)+
    geom_bar(position="fill", stat="identity")+
   scale_fill_brewer(palette = 'Set2')+
  theme_classic()+
  ylab('Percent abundance of compounds')+
  xlab('Fungal isolates')+
  labs(title= "Differentially expressed compounds by compound class",
       comment="Differetally expressed compounds pval.adj< 0.05 ")
stack.class
figure_file <- file.path(figures_dir, 'Class.Percent.abund.png')
ggsave(figure_file, stack.class, dpi=400)

stack.class2<-ggplot(sig.data.all, aes(fill=Class, y=n, x=Fungi)) + 
  facet_wrap(~Comment)+
    geom_bar(position="stack", stat="identity")+
  scale_fill_brewer(palette = 'Set2')+
  ylab('Number of compounds')+
  xlab('Fungal isolates')+
  labs(title= "Differentially expressed compounds by compound class",
       caption="Differetally expressed compounds pval< 0.05 ")+
  theme_classic()
stack.class2
figure_file <- file.path(figures_dir, 'Class.Relative.abund.png')
ggsave(figure_file, stack.class2, dpi=400)
```
```

```{r warning=FALSE}

figure_file <- file.path(figures_dir, 'All_identified_features_heatmap.pdf')

pdf(figure_file, width = 15, height = 10)
pheatmap(named_compounds.matrix,
         clustering_distance_rows = 'correlation',
         clustering_distance_cols = 'correlation',
         scale = 'row',
         annotation_col = col_annot,
         annotation_colors = annot_colors,
         color = mapcolor,
         cutree_cols = 4,
         cutree_rows = 5,
         fontsize_row = 6,
         main = 'Identified features (scaled AUC)'
)
dev.off()
```

[Scaled Area under the curve (AUC) heatmap for only features that were identified](`r toString(figure_file)`)

## 4.3 KEGG Up and Down Regulated 
```{r}

F1246.sig$KEGGPathways<-as.factor(F1246.sig$"KEGG Pathways")  
F1246.count<- F1246.sig%>% 
  select(FeatureID, KEGGPathways, Comment) %>% 
  separate_rows(KEGGPathways, sep = ';') %>% 
  group_by(KEGGPathways, Comment ) %>% 
  count(KEGGPathways) %>%
   drop_na(KEGGPathways)%>% filter(Comment != "Only present in control")
  
class.plot<- ggplot(F1246.count)+
geom_bar(aes(y=KEGGPathways, x=n,fill= Comment), stat="identity", alpha=0.7)+
  #scale_fill_manual(values = c("lightskyblue","firebrick2"))+
  theme_classic()+xlab("Number of Factors")+ylab("KEGG Pathways")+
  labs(title="Up and down regulated KEGG pathways of Isolate 1246")
class.plot

figure_file <- file.path(figures_dir, 'KEGGrank_F1246.png')
ggsave(figure_file, class.plot, width=15, height=10, unit='cm')


F1177.sig$KEGGPathways<-as.factor(F1177.sig$"KEGG Pathways")  
F1177.count<- F1177.sig%>% 
  select(FeatureID, KEGGPathways, Comment) %>% 
  separate_rows(KEGGPathways, sep = ';') %>% 
  group_by(KEGGPathways, Comment ) %>% 
  count(KEGGPathways) %>%
   drop_na(KEGGPathways)%>% filter(Comment != "Only present in control")
  
class.plot<- ggplot(F1177.count)+
geom_bar(aes(y=KEGGPathways, x=n,fill= Comment), stat="identity", alpha=0.7)+
  #scale_fill_manual(values = c("lightskyblue","firebrick2"))+
  theme_classic()+xlab("Number of Factors")+ylab("KEGG Pathways")+
  labs(title="Up and down regulated KEGG pathways of Isolate 1177")
class.plot

figure_file <- file.path(figures_dir, 'KEGGrank_F1177.png')
ggsave(figure_file, class.plot, width=15, height=10, unit='cm')


F1154.sig$KEGGPathways<-as.factor(F1154.sig$"KEGG Pathways")  
F1154.count<- F1154.sig%>% 
  select(FeatureID, KEGGPathways, Comment) %>% 
  separate_rows(KEGGPathways, sep = ';') %>% 
  group_by(KEGGPathways, Comment ) %>% 
  count(KEGGPathways) %>%
   drop_na(KEGGPathways)%>% filter(Comment != "Only present in control")
  
class.plot<- ggplot(F1154.count)+
geom_bar(aes(y=KEGGPathways, x=n,fill= Comment), stat="identity", alpha=0.7)+
  #scale_fill_manual(values = c("lightskyblue","firebrick2"))+
  theme_classic()+xlab("Number of Factors")+ylab("KEGG Pathways")+
  labs(title="Up and down regulated KEGG pathways of Isolate 1154")
class.plot

figure_file <- file.path(figures_dir, 'KEGGrank_F1154.png')
ggsave(figure_file, class.plot, width=15, height=10, unit='cm')
```




## 4.38 Upregulated Secondary Metabolites
```{r}

F1246.sig$KEGGPathways<-as.factor(F1246.sig$"KEGG Pathways")  
F1246.reg<- F1246.sig %>% 
  #select(FeatureID, KEGGPathways, Comment) %>% 
  separate_rows(KEGGPathways, sep = ';') %>% 
  filter(KEGGPathways== 'Biosynthesis of secondary metabolites')%>% 
 filter(Comment == "Upregulated")

table_file <- file.path(tables_dir, 'Upreg.SecondaryMetabolites_F1246.csv')
write_csv( F1246.reg, table_file)
  
F1177.sig$KEGGPathways<-as.factor(F1177.sig$"KEGG Pathways")  
F1177.reg<- F1177.sig%>% 
  #select(FeatureID, KEGGPathways, Comment) %>% 
  separate_rows(KEGGPathways, sep = ';') %>% 
  filter(KEGGPathways== 'Biosynthesis of secondary metabolites')%>% 
 filter(Comment == "Upregulated")
table_file <- file.path(tables_dir, 'Upreg.SecondaryMetabolites_F1177.csv')
write_csv( F1177.reg, table_file)

F1154.sig$KEGGPathways<-as.factor(F1154.sig$"KEGG Pathways")  
F1154.reg<- F1154.sig%>% 
  #select(FeatureID, KEGGPathways, Comment) %>% 
  separate_rows(KEGGPathways, sep = ';') %>% 
  filter(KEGGPathways== 'Biosynthesis of secondary metabolites')%>% 
 filter(Comment == "Upregulated")
table_file <- file.path(tables_dir, 'Upreg.SecondaryMetabolites_F1154.csv')
write_csv( F1154.reg, table_file)
```

### All dysregulated features

```{r}
# Create an AUC matrix that contains only features that are differentially expressed

sig_features.matrix <- norm.matrix[row.names(norm.matrix) %in% sig_features, ]

row.names(sig_features.matrix) <- compounds_table$name4plot[match(row.names(sig_features.matrix), compounds_table$FeatureID)]
```



[Scaled Area under the curve (AUC) heatmap for all dysregulated features](`r toString(figure_file)`)

### Dysregulated features that were identified

```{r }
named_sig_features.matrix <- sig_features.matrix[rownames(sig_features.matrix) %in% named_compounds$Name, ]

figure_file <- file.path(figures_dir, 'Dysreg_identified_features.pdf')

pdf(figure_file, width = 10)
pheatmap(named_sig_features.matrix,
         clustering_distance_rows = 'correlation',
         clustering_distance_cols = 'correlation',
         scale = 'row',
         annotation_col = col_annot,
         annotation_colors = annot_colors,
         color = mapcolor,
         cutree_cols = 4,
         fontsize_row = 8,
