---
title: "DifferentialAnalysis"
output: html_notebook
---

---
title: "RP Differential Analysis All Isolates+Seeds All"
output: html_notebook
---

# 1. Importing Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggrepel)
library(ggsci)
library(pheatmap)
library(RColorBrewer)
library(readxl)
library(UpSetR)
source('functions_cdis_diff.R')
```

Change to what data you are using
```{r}
data= 'RP'
```

# 2. Import data


Because differential analysis includes the calculations of means, the data to be used in this section is the normalized untransformed data

```{r set_path, message=FALSE, warning=FALSE}
# set path variables
project_dir <- getwd()
project_name <- 'RP_All_1154'
figures_dir <- file.path(project_dir, paste0(project_name, '_output_figures'))
tables_dir <- file.path(project_dir,  paste0(project_name, '_output_tables'))
norm_auc_table_file <- file.path(tables_dir, 'normalized_untransformed_auc_table.csv')
norm_auc_table_nogap_file <- file.path(tables_dir, 'normalized_untransformed_auc_table_nogap.csv')
# If the flag was setted before, nothing needs to be changed here, the correct file will be automatically used

  compounds_table_file <- file.path(tables_dir, 'compounds_table.csv')


# Load auc_table

norm.matrix <- read_csv(norm_auc_table_nogap_file)
norm.matrix <- column_to_rownames(norm.matrix, var = '...1')

norm.matrix_nogap <- read_csv(norm_auc_table_nogap_file)
norm.matrix_nogap <- column_to_rownames(norm.matrix_nogap, var = '...1')

# Load compounds table and add a column that has names for identified compounds and FeatureID for the rest (for plotting puRPouses)

compounds_table <- read_csv(compounds_table_file)

# Import metadata and fix names
metadata_file <- file.path(tables_dir, 'fixed_metadata.csv')
metadata <- read_csv(metadata_file)



```



# 3. Calculate ratios and log2fold-change

## 3.1 Get samples per each treatment

```{r}

# Get the average AUC per each of the treatments
## Get samples per treatment

ERLE.samples <- get_samples(metadata, Treatment = 'Plant', value = 'ERLE')
ERIN.samples <- get_samples(metadata, Treatment = 'Plant', value = 'ERIN')
BOCU.samples <- get_samples(metadata, Treatment = 'Plant', value = 'BOCU')
LEDU.samples <- get_samples(metadata, Treatment = 'Plant', value = 'LEDU')
CTRL.samples <- get_samples(metadata, Treatment = 'Fungi', value = 'CTRL')

```

## 3.2 Get table of differentially expressed features per each comparison

The following function will calculate ratio, log2FC, p values and adjusted pvalues. If no replicates are available for EACH treatment please use the get_diff_table_no_pval() function

```{r}

# The following function will calculate ratio, log2FC, p values and adjusted pvalues. If no replicates are available for EACH treatment
# please use the get_diff_table_no_pval() function

ERLE_to_CTRL.diff_table <- get_diff_table(norm.matrix, treatment.sample_list = ERLE.samples, control.sample_list = CTRL.samples, log2_transformed = TRUE)
ERIN_to_CTRL.diff_table <- get_diff_table(norm.matrix, treatment.sample_list = ERIN.samples, control.sample_list = CTRL.samples, log2_transformed = TRUE)
BOCU_to_CTRL.diff_table <- get_diff_table(norm.matrix, treatment.sample_list = BOCU.samples, control.sample_list = CTRL.samples, log2_transformed = TRUE)
LEDU_to_CTRL.diff_table <- get_diff_table(norm.matrix, treatment.sample_list = LEDU.samples, control.sample_list = CTRL.samples, log2_transformed = TRUE)
```

Merge differentially expressed features with the annotation

```{r}

# Create dataframes for the up and downregulated metabolites at each time point and merge them with the compound information

ERLE_to_CTRL.diff_table <- compounds_table %>% 
  select( - contains('Results'), -SampleID, -AUC,  -Fungi, - Plant) %>% 
  right_join(ERLE_to_CTRL.diff_table, by = 'FeatureID') %>% 
  distinct() %>% 
  filter(!is.nan(log2FC)) %>% 
  mutate(Comment = case_when(log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

ERLE_to_CTRL.diff_table$Comment <- factor(ERLE_to_CTRL.diff_table$Comment, 
                               levels = c( 'Downregulated', 'Upregulated' ))

table_file <- file.path(tables_dir, 'Diff_expressed_ERLE.csv')
write_csv(ERLE_to_CTRL.diff_table, table_file )

ERIN_to_CTRL.diff_table<- compounds_table %>% 
  select(-contains('Annotation Source:'), - contains('Results'), -SampleID, -AUC, -Plant, - Fungi) %>%
  right_join(ERIN_to_CTRL.diff_table, by = 'FeatureID') %>% 
  distinct() %>% 
  filter(!is.nan(log2FC)) %>% 
  mutate(Comment = case_when(log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

ERIN_to_CTRL.diff_table$Comment <- factor(ERIN_to_CTRL.diff_table$Comment, 
                               levels = c( 'Downregulated', 'Upregulated'))

table_file <- file.path(tables_dir, 'Diff_expressed_ERIN.csv')
write_csv(ERIN_to_CTRL.diff_table, table_file )

BOCU_to_CTRL.diff_table <- compounds_table %>% 
  select(-contains('Annotation Source:'), - contains('Results'), -SampleID, -AUC, -Plant, -Fungi) %>% 
  right_join(BOCU_to_CTRL.diff_table, by = 'FeatureID') %>% 
  distinct() %>% 
  filter(!is.nan(log2FC)) %>% 
  mutate(Comment = case_when(log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

BOCU_to_CTRL.diff_table$Comment <- factor(BOCU_to_CTRL.diff_table$Comment, 
                               levels = c('Downregulated', 'Upregulated' ))

table_file <- file.path(tables_dir, 'Diff_expressed_BOCU.csv')
write_csv(BOCU_to_CTRL.diff_table, table_file )

LEDU_to_CTRL.diff_table <- compounds_table %>% 
  select(-contains('Annotation Source:'), - contains('Results'), -SampleID, -AUC, -Plant, -Fungi) %>% 
  right_join(LEDU_to_CTRL.diff_table, by = 'FeatureID') %>% 
  distinct() %>% 
  filter(!is.nan(log2FC)) %>% 
  mutate(Comment = case_when(log2FC < 0 ~ 'Downregulated',
                             log2FC > 0 ~ 'Upregulated'))

LEDU_to_CTRL.diff_table$Comment <- factor(LEDU_to_CTRL.diff_table$Comment, 
                               levels = c('Downregulated', 'Upregulated' ))

table_file <- file.path(tables_dir, 'Diff_expressed_LEDU.csv')
write_csv(LEDU_to_CTRL.diff_table, table_file )

```

Extract most differentially expressed features

```{r warning=FALSE}

# Extract significant features of each comparison based on adjusted pvalue
sig_features <- c(ERLE_to_CTRL.diff_table$FeatureID[ERLE_to_CTRL.diff_table$pval < 0.05],
                  ERIN_to_CTRL.diff_table$FeatureID[ERIN_to_CTRL.diff_table$pval < 0.05],
                  BOCU_to_CTRL.diff_table$FeatureID[BOCU_to_CTRL.diff_table$pval < 0.05],

LEDU_to_CTRL.diff_table$FeatureID[LEDU_to_CTRL.diff_table$pval < 0.05])

sig_features <- unique(sig_features)

```


# 4. Plots of dysregulated features

## 4.1 Number of dysregulated features in each group

```{r}
# removed pval adj to regular pval
num_diff_features <- tibble(Comparison = rep(c('ERLE_to_CTRL', 'ERIN_to_CTRL', 'BOCU_to_CTRL', 'LEDU_to_CTRL'), each = 2),
                            Type = rep(c('Name', 'No name'), 4),
                            count = 0)
num_diff_features$Type <- factor(num_diff_features$Type, levels = c('No name', 'Name'))

num_diff_features[1, 3] <- sum(ERLE_to_CTRL.diff_table$pval < 0.05 & !is.na(ERLE_to_CTRL.diff_table$Name), na.rm = TRUE)
num_diff_features[2, 3] <- sum(ERLE_to_CTRL.diff_table$pval < 0.05 & is.na(ERLE_to_CTRL.diff_table$Name), na.rm = TRUE)
num_diff_features[3, 3] <- sum(ERIN_to_CTRL.diff_table$pval < 0.05 & !is.na(ERIN_to_CTRL.diff_table$Name), na.rm = TRUE)
num_diff_features[4, 3] <- sum(ERIN_to_CTRL.diff_table$pval< 0.05 & is.na(ERIN_to_CTRL.diff_table$Name), na.rm = TRUE)
num_diff_features[5, 3] <- sum(BOCU_to_CTRL.diff_table$pval < 0.05 & !is.na(BOCU_to_CTRL.diff_table$Name), na.rm = TRUE)
num_diff_features[6, 3] <- sum(BOCU_to_CTRL.diff_table$pval< 0.05 & is.na(BOCU_to_CTRL.diff_table$Name), na.rm = TRUE)
num_diff_features[5, 3] <- sum(LEDU_to_CTRL.diff_table$pval< 0.05 & !is.na(LEDU_to_CTRL.diff_table$Name), na.rm = TRUE)
num_diff_features[6, 3] <- sum(LEDU_to_CTRL.diff_table$pval < 0.05 & is.na(LEDU_to_CTRL.diff_table$Name), na.rm = TRUE)

num_diff_plot <- ggplot(num_diff_features)+
  geom_bar(aes(y=count, x=Comparison,fill= Type), stat="identity", alpha=0.7)+
  theme_classic()+xlab("Comparison")+ylab("Count")
  
num_diff_plot
# shows snumber of significantly regulated compounds
figure_file <- file.path(figures_dir, 'Num_diff_features.png')
ggsave(figure_file, num_diff_plot, dpi = 300)
```

## 4.2 Dysregulated features shared at different timepoints

An **upset plot**is used instead of a Venn Diagram as it provides a better visualization of the number of figures that are being shared among groups

```{r}
#Filter diff tables by pval.adj

ERLE.sig<-ERLE_to_CTRL.diff_table%>% 
  filter(pval<0.05)
ERIN.sig<-ERIN_to_CTRL.diff_table%>% 
  filter(pval<0.05)
BOCU.sig<-BOCU_to_CTRL.diff_table%>% 
  filter(pval<0.05)
LEDU.sig<-LEDU_to_CTRL.diff_table%>% 
  filter(pval<0.05)

#Get list of features at each time point for each of the cases
ERLE.upregulated<-get_vectors(ERLE.sig, 'Comment', 'Upregulated', 'FeatureID')
ERLE.downregulated<-get_vectors(ERLE.sig, 'Comment', 'Downregulated', 'FeatureID')
#ERLE.not_in_control<-get_vectors(ERLE.sig, 'Comment', 'Not #present in control', 'FeatureID')
#ERLE.only_in_control<-get_vectors(ERLE.sig, 'Comment', 'Only #present in control', 'FeatureID')

ERIN.upregulated<-get_vectors(ERIN.sig, 'Comment', 'Upregulated', 'FeatureID')
ERIN.downregulated<-get_vectors(ERIN.sig, 'Comment', 'Downregulated', 'FeatureID')
#ERIN.not_in_control<-get_vectors(ERIN.sig, 'Comment', 'Not present in control', 'FeatureID')
#ERIN.only_in_control<-get_vectors(ERIN.sig, 'Comment', 'Only present in control', 'FeatureID')

BOCU.upregulated<-get_vectors(BOCU.sig, 'Comment', 'Upregulated', 'FeatureID')
BOCU.downregulated<-get_vectors(BOCU.sig, 'Comment', 'Downregulated', 'FeatureID')
#BOCU.not_in_control<-get_vectors(BOCU.sig, 'Comment', 'Not present in control', 'FeatureID')
#BOCU.only_in_control<-get_vectors(BOCU.sig, 'Comment', 'Only present in control', 'FeatureID')

LEDU.upregulated<-get_vectors(LEDU.sig, 'Comment', 'Upregulated', 'FeatureID')
LEDU.downregulated<-get_vectors(LEDU.sig, 'Comment', 'Downregulated', 'FeatureID')
#LEDU.not_in_control<-get_vectors(LEDU.sig, 'Comment', 'Not present in control', 'FeatureID')
#LEDU.only_in_control<-get_vectors(LEDU.sig, 'Comment', 'Only present in control', 'FeatureID')
```
```{r fig.height=8}
upset_input<-list(ERLE.upregulated=ERLE.upregulated,
                    ERLE.downregulated=ERLE.downregulated,
                    #ERLE.not_in_control=ERLE.not_in_control,
                    #ERLE.only_in_control=ERLE.only_in_control,
                    ERIN.upregulated=ERIN.upregulated,
                    ERIN.downregulated=ERIN.downregulated,
                    #ERIN.not_in_control=ERIN.not_in_control,
                    #ERIN.only_in_control=ERIN.only_in_control,
                    BOCU.upregulated=BOCU.upregulated,
                    BOCU.downregulated=BOCU.downregulated,
                  LEDU.upregulated=LEDU.upregulated,
                  LEDU.downregulated=LEDU.downregulated)
                    #BOCU.not_in_control=BOCU.not_in_control,
                    #BOCU.only_in_control=BOCU.only_in_control)

upset_metadata<-data.frame(sets=as.vector(names(upset_input)),
                         Time=rep(c('ERLE', 'ERIN', 'BOCU', 'LEDU'), each=4))

upset_features<-upset(fromList(upset_input), order.by="freq", cutoff=1, nsets=12,
                    mainbar.y.label='Number of shared features', sets.x.label='Number of features',
                    text.scale=c(1.5, 1, 1.5, 1, 1.3, 1.3),
                    set.metadata=list(data=upset_metadata, 
                                        plots=list(list(type='matrix_rows', column='Time', 
                                                          colors=c(ERLE='#EFC000FF', ERIN='#868686FF', BOCU='#CD534CFF')))))
upset_features

figure_file<-file.path(figures_dir, 'Upset-KEGG.png')
png(figure_file, width=800, height=800, res=100)
upset_features
dev.off()
```

## 4.2 Volcano plots

labels if L2FC is doubble lfc.t

```{r warning=FALSE}

lfc.t <- 2
pval.t <- 0.05

ERLE_to_CTRL_volcano <- plot_volcano(ERLE_to_CTRL.diff_table, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'Isolate ERLE vs Media')
ERLE_to_CTRL_volcano

figure_file <- file.path(figures_dir, 'ERLE_CTRL_volcano.png')
ggsave(figure_file, ERLE_to_CTRL_volcano, width=20, height=20, unit='cm')


ERIN_to_CTRL_volcano <- plot_volcano(ERIN_to_CTRL.diff_table, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'Isolate ERIN vs Media')
ERIN_to_CTRL_volcano

figure_file <- file.path(figures_dir, 'ERIN_CTRL_volcano.png')
ggsave(figure_file, ERIN_to_CTRL_volcano, width=20, height=20, unit='cm')

BOCU_to_CTRL_volcano <- plot_volcano(BOCU_to_CTRL.diff_table, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'Isolate BOCU vs Media')
BOCU_to_CTRL_volcano

figure_file <- file.path(figures_dir, 'BOCU_CTRL_volcano.png')
ggsave(figure_file, BOCU_to_CTRL_volcano, width=20, height=20, unit='cm')

LEDU_to_CTRL_volcano <- plot_volcano(LEDU_to_CTRL.diff_table, log2FC, pval.adj, lfc.t, pval.t) +
  labs(subtitle = 'Isolate LEDU vs Media')
LEDU_to_CTRL_volcano

figure_file <- file.path(figures_dir, 'LEDU_CTRL_volcano.png')
ggsave(figure_file, LEDU_to_CTRL_volcano, width=20, height=20, unit='cm')
```


```{r}
# Filter diff tables by pval.adj
BOCU.sig <- BOCU_to_CTRL.diff_table %>% 
  filter(pval< 0.05)
ERLE.sig <- ERLE_to_CTRL.diff_table %>% 
  filter(pval < 0.05)
ERIN.sig <- ERIN_to_CTRL.diff_table %>% 
  filter(pval < 0.05)
LEDU.sig<-LEDU_to_CTRL.diff_table%>%
  filter(pval<0.5)

BOCU.sig2 <- BOCU.sig %>% 
  filter(log2FC > 2 | log2FC < 0.5)
ERLE.sig2 <- ERLE.sig %>% 
  filter(log2FC > 2 | log2FC < 0.5)
ERIN.sig2 <- ERIN.sig %>% 
  filter(log2FC > 2 | log2FC < 0.5)
LEDU.sig2 <- LEDU.sig %>% 
  filter(log2FC > 2 | log2FC < 0.5)

ERLE.sig2$isolate<- "ERLE"
ERIN.sig2$isolate<- "ERIN"
BOCU.sig2$isolate<- "BOCU"
LEDU.sig2$isolate<-"LEDU"
sig.all<- rbind(ERLE.sig2, ERIN.sig2, BOCU.sig2, LEDU.sig2)

table_file <- file.path(tables_dir, 'Sig.diff.expressed.allIsolates_L2FC_0.3.csv')
write_csv(sig.all, table_file )
# Get list of features at each time point for each of the cases
BOCU.upregulated <- get_vectors(BOCU.sig, 'Comment', 'Upregulated', 'FeatureID')
BOCU.downregulated <- get_vectors(BOCU.sig, 'Comment', 'Downregulated', 'FeatureID')
#BOCU.not_in_control <- get_vectors(BOCU.sig, 'Comment', 'Not present in control', 'FeatureID')
#BOCU.only_in_control <- get_vectors(BOCU.sig, 'Comment', 'Only present in control', 'FeatureID')
ERLE.upregulated <- get_vectors(ERLE.sig, 'Comment', 'Upregulated', 'FeatureID')
ERLE.downregulated <- get_vectors(ERLE.sig, 'Comment', 'Downregulated', 'FeatureID')
#ERLE.not_in_control <- get_vectors(ERLE.sig, 'Comment', 'Not present in control', 'FeatureID')
#ERLE.only_in_control <- get_vectors(ERLE.sig, 'Comment', 'Only present in control', 'FeatureID')
ERIN.upregulated <- get_vectors(ERIN.sig, 'Comment', 'Upregulated', 'FeatureID')
ERIN.downregulated <- get_vectors(ERIN.sig, 'Comment', 'Downregulated', 'FeatureID')
#ERIN.not_in_control <- get_vectors(ERIN.sig, 'Comment', 'Not present in control', 'FeatureID')
#ERIN.only_in_control <- get_vectors(ERIN.sig, 'Comment', 'Only present in control', 'FeatureID')
```
## 4.3 Van Krevelen Diagrams of dysregulated features

**Dysregulated features at ERLE**

```{r}

ERLE_vank<-plot_vank(ERLE.sig, Class, Comment)
ERLE_vank

figure_file<-file.path(figures_dir, 'DE_vank_ERLE.png')
ggsave(figure_file, ERLE_vank)
```

**Dysregulated features at ERIN*

```{r}
ERIN_vank<-plot_vank(ERIN.sig, Class, Comment)
ERIN_vank

figure_file<-file.path(figures_dir, 'DE_vank_ERIN.png')
ggsave(figure_file, ERIN_vank)
```

**Dysregulated features at BOCU**

```{r}
BOCU_vank<-plot_vank(BOCU.sig, Class, Comment)
BOCU_vank

figure_file<-file.path(figures_dir, 'DE_vank_BOCU.png')
ggsave(figure_file, BOCU_vank)
```

Van Krevelen of all

```{r}

sig_all<-rbind(ERLE.sig%>% select(FeatureID, Comment) %>% mutate(Plant='ERLE'),
                 ERIN.sig%>% select(FeatureID, Comment) %>% mutate(Plant='ERIN'),
                 BOCU.sig%>% select(FeatureID, Comment) %>% mutate(Plant='BOCU'),
               LEDU.sig%>% select(FeatureID, Comment) %>% mutate(Plant='LEDU'))

sig_all_compounds_table<-compounds_table%>% 
  select(FeatureID, H_to_C, O_to_C, Class, Fungi, GFE, Plant) %>% 
  distinct() %>% 
  left_join(sig_all, by=c('FeatureID', 'Plant'))%>%
  na.omit()



my_colors=c('Upregulated'='tomato2',
              'Downregulated'='lightsteelblue')



dysreg_vk<-ggplot(sig_all_compounds_table,
                    aes(x=O_to_C,
                        y=H_to_C,
                        color=Class)) +
  geom_point() +
  scale_color_manual(values=my_colors) +
  theme_classic() +
  labs(x='O:C',
       y='H:C',
       title='Van Krevelen Diagram ',
       color='Assigned compound class') +
  #new_scale_color() +
  #class_rect+
  scale_color_manual(values=get_palette('d3', 8)) +
  theme(plot.title=element_text(face='bold', hjust=0.5)) +
  facet_wrap(~Fungi*Comment, 3)

dysreg_vk

figure_file<-file.path(figures_dir, 'dysreg_vk.png')
ggsave(figure_file, dysreg_vk)
```

## 4.4 GFE plots of dysregulated features

**Dysregulated features at ERLE**

```{r}

my_colors=c( 
              'Upregulated'='tomato2',
              'Downregulated'='lightsteelblue')

ERLE_GFE_box<-plot_boxplot(ERLE.sig, Comment, GFE, Comment, my_colors)+ labs(title='Isolate 1154')
ERLE_GFE_box

figure_file<-file.path(figures_dir, 'DE_GFE_ERLE.png')
ggsave(figure_file, ERLE_GFE_box)
```

**Dysregulated features at ERIN**

```{r}
ERIN_GFE_box<-plot_boxplot(ERIN.sig, Comment, GFE, Comment, my_colors)+ labs(title='Isolate 1177')
ERIN_GFE_box

figure_file<-file.path(figures_dir, 'DE_GFE_ERIN.png')
ggsave(figure_file, ERIN_GFE_box)
```

**Dysregulated features at BOCU**

```{r}
BOCU_GFE_box<-plot_boxplot(BOCU.sig, Comment, GFE, Comment, my_colors)+ labs(title= 'Isolate 1246')
BOCU_GFE_box

figure_file<-file.path(figures_dir, 'DE_GFE_BOCU.png')
ggsave(figure_file, BOCU_GFE_box)


all_GFE_box<- ggplot(sig_all_compounds_table,
         aes(x = Comment,
             y = GFE,
             fill = Comment)) +
    geom_violin() +
    geom_boxplot(width=0.1, color="black", alpha=0.2) +
    scale_fill_manual(values = my_colors) +
    theme_classic() +
    theme(plot.title = element_text(face = 'bold', hjust = 0.5))+
  facet_wrap(~Plant)
all_GFE_box
figure_file<-file.path(figures_dir, 'all_GFE.png')
ggsave(figure_file, all_GFE_box)

```




```{r}

BOCU.count<- BOCU.sig%>% #BOCU.sig only includes sig. up/down regulated features 
    select(FeatureID, Class, Comment, pval.adj)%>%
    count(Class, Comment)

BOCU.count <- within(BOCU.count, n[Comment=="Downregulated"] <- (n*(-1))) #Makes downregulated features negative in the figure

  
class.plot<- ggplot(BOCU.count)+
geom_bar(aes(y=Class, x=n,fill= Comment), stat="identity", alpha=0.7)+
   scale_fill_manual(values = c("lightsteelblue","tomato2"))+
  theme_classic()+xlab("Number of Features")+ylab("Class")+
  labs(title="Class ranks of Isolate 1246")
class.plot

figure_file <- file.path(figures_dir, 'Class_rank_BOCU.png')
ggsave(figure_file, class.plot, width=15, height=10, unit='cm')

# I then repeated theses scripts for each isolate 


ERIN.count<- ERIN.sig%>% 
    select(FeatureID, Class, Comment, pval.adj)%>%
    count(Class, Comment)
ERIN.count <- within(ERIN.count, n[Comment=="Downregulated"] <- (n*(-1)))
    #filter(Comment != "Only present in control")
  
class.plot<- ggplot(ERIN.count)+
geom_bar(aes(y=Class, x=n,fill= Comment), stat="identity", alpha=0.7)+
   scale_fill_manual(values = c("lightsteelblue","tomato2"))+
  theme_classic()+xlab("Number of Features")+ylab("Class")+
  labs(title="Class ranks of Isolate 1177")
class.plot

figure_file <- file.path(figures_dir, 'Class_rank_ERIN.png')
ggsave(figure_file, class.plot, width=15, height=10, unit='cm')

ERLE.count<- ERLE.sig%>% 
    select(FeatureID, Class, Comment, pval.adj, pval)%>%
    count(Class, Comment)
ERLE.count <- within(ERLE.count, n[Comment=="Downregulated"] <- (n*(-1)))
    #filter(Comment != "Only present in control")
  
class.plot<- ggplot(ERLE.count)+
geom_bar(aes(y=Class, x=n,fill= Comment), stat="identity", alpha=0.7)+
   scale_fill_manual(values = c("lightsteelblue","tomato2"))+
  theme_classic()+xlab("Number of Features")+ylab("Class")+
  labs(title="Class ranks of Isolate 1154")
class.plot

figure_file <- file.path(figures_dir, 'Class_rank_ERLE.png')
ggsave(figure_file, class.plot, width=15, height=10, unit='cm')

ERLE.count$isolate<- 1154
ERIN.count$isolate<- 1177
BOCU.count$isolate<- 1246
class.rank.all<- rbind(ERLE.count, ERIN.count, BOCU.count)


class.plot<- ggplot(class.rank.all)+
geom_bar(aes(y=Class, x=n,fill= Comment), stat="identity", alpha=0.7)+
  facet_wrap(~isolate, 3)+
   scale_fill_manual(values = c("lightsteelblue","tomato2"))+
  theme_classic()+xlab("Number of Features")+ylab("Class")+
  labs(title="Class ranks")
class.plot

figure_file <- file.path(figures_dir, 'Class_rank_all.png')
ggsave(figure_file, class.plot, width=15, height=10, unit='cm')

class.plot<- ggplot(class.rank.all)+
geom_bar(aes(y=Class, x=n,fill= Comment), stat="identity", alpha=0.7)+
  facet_wrap(~isolate, 1)+
   scale_fill_manual(values = c("lightsteelblue","tomato2"))+
  theme_classic()+xlab("Number of Features")+ylab("Class")+
  labs(title="Class ranks")
class.plot

figure_file <- file.path(figures_dir, 'Class_rank_all_2.png')
ggsave(figure_file, class.plot, width=15, height=10, unit='cm')



```






# 5. Hierarchical clustering analysis using heatmaps

Multiple heatmaps will be plotted, based on the differential expression and if the compounds were or not assigned and structure

### Normalized Area under the curve (AUC) heatmap for all detected features

```{r warning=FALSE}

# Initialize graphical device
#dev.off()

# Set sampleID as row.names to annotate heatmap

col_annot <- column_to_rownames(metadata, var = 'SampleID') 
mapcolor <- colorRampPalette(brewer.pal(11, 'RdYlBu'))(100)[100:1]

figure_file <- file.path(figures_dir, 'All_features_heatmap.pdf')

annot_colors <- list(
  Time = c(CTRL ='#0073C2FF', ERLE = '#EFC000FF', ERIN = '#868686FF', BOCU = '#CD534CFF'),
  Material = c(Litter = 'green4', Litter_Peat = 'chocolate4'),
  Origin = c(LC_MS2 = '#0086cb', NMR = '#d23936'))

pdf(figure_file)
pheatmap(norm.matrix_nogap,
         clustering_distance_rows = 'correlation',
         clustering_distance_cols = 'correlation',
         scale = 'row',
         annotation_col = col_annot,
         annotation_colors = annot_colors,
         color = mapcolor,
         show_rownames = FALSE,
         cutree_cols = 5,
         main = 'All features (scaled AUC)'
)
#dev.off()
```

[Scaled Area under the curve (AUC) heatmap for all detected features](`r toString(figure_file)`)

### Normalized Area under the curve (AUC) heatmap for only features that were identified

```{r}
# Extract all features that have names
named_compounds <- compounds_table %>% 
  select(FeatureID, Name, name4plot) %>% 
  filter(!is.na(Name)) %>% 
  distinct()

# Create a matrix of only identified features

named_compounds.matrix <- norm.matrix[rownames(norm.matrix) %in% named_compounds$FeatureID,]
row.names(named_compounds.matrix) <- compounds_table$name4plot[match(row.names(named_compounds.matrix), compounds_table$FeatureID)]

```


```{r}

ERLE.sig$Plant<- 'ERLE'
ERIN.sig$Plant<- 'ERIN'
BOCU.sig$Plant<- 'BOCU'
LEDU.sig$Plant<-'LEDU'


#sig.data.all<- rbind(ERLE.sig, ERIN.sig, BOCU.sig, LEDU.sig)
#compounds_table$Class <- gsub( "Lignin", "Phenol", compounds_table$Class)
sig_all_compounds_table$Class <- gsub("Tannin","Other",  sig_all_compounds_table$Class)

sig.data.all<- sig_all_compounds_table %>%
  mutate(class = case_when(
    endsWith(Class, "Lignin") ~ "Phenol",
    endsWith(Class, "Tannin") ~ "Highly oxygenated compounds"
  ))%>%
  select(FeatureID, Class, Plant, Comment)%>%
  count(Class, Plant, Comment)
  


stack.class<-ggplot(sig.data.all, aes(fill=Class, y=n, x=Plant)) + 
    facet_wrap(~Comment)+
    geom_bar(position="fill", stat="identity")+
   scale_fill_brewer(palette = 'Set2')+
  theme_classic()+
  ylab('Percent abundance of compounds')+
  xlab('Fungal isolates')+
  labs(title= "Differentially expressed compounds by compound class",
       comment="Differetally expressed compounds pval.adj< 0.05 ")
stack.class
figure_file <- file.path(figures_dir, 'Class.Percent.abund.png')
ggsave(figure_file, stack.class, dpi=200)

stack.class2<-ggplot(sig.data.all, aes(fill=Class, y=n, x=Plant)) + 
  facet_wrap(~Comment)+
    geom_bar(position="stack", stat="identity")+
  scale_fill_brewer(palette = 'Set2')+
  ylab('Number of compounds')+
  xlab('Fungal isolates')+
  labs(title= "Differentially expressed compounds by compound class",
       caption="Differetally expressed compounds pval< 0.05 ")+
  theme_classic()
stack.class2
figure_file <- file.path(figures_dir, 'Class.Relative.abund.png')
ggsave(figure_file, stack.class2, dpi=200)
```


```{r}
figure_file <- file.path(figures_dir, 'All_identified_features_heatmap.pdf')

pdf(figure_file, width = 15, height = 10)
pheatmap(named_compounds.matrix,
         clustering_distance_rows = 'correlation',
         clustering_distance_cols = 'correlation',
         scale = 'row',
         annotation_col = col_annot,
         annotation_colors = annot_colors,
         color = mapcolor,
         cutree_cols = 4,
         cutree_rows = 5,
         fontsize_row = 6,
         main = 'Identified features (scaled AUC)'
)
dev.off()
```

[Scaled Area under the curve (AUC) heatmap for only features that were identified](`r toString(figure_file)`)

## 4.3 KEGG Up and Down Regulated 
```{r}

BOCU.sig$KEGGPathways<-as.factor(BOCU.sig$"KEGG Pathways")  

BOCU.count<- BOCU.sig%>% 
  select(FeatureID, KEGGPathways, Comment) %>% 
  separate_rows(KEGGPathways, sep = ';') %>% 
  group_by(KEGGPathways, Comment ) %>% 
  count(KEGGPathways) %>%
   drop_na(KEGGPathways)%>% filter(Comment != "Only present in control")
  BOCU.count <- within(BOCU.count, n[Comment=="Downregulated"] <- (n*(-1)))
class.plot<- ggplot(BOCU.count)+
geom_bar(aes(y=KEGGPathways, x=n,fill= Comment), stat="identity", alpha=0.7)+
   scale_fill_manual(values = c("lightsteelblue","tomato2"))+
  theme_classic()+xlab("Number of Features")+ylab("KEGG Pathways")+
  labs(title="Up and down regulated KEGG pathways of Isolate 1246")
class.plot

figure_file <- file.path(figures_dir, 'KEGGrank_BOCU.png')
ggsave(figure_file, class.plot, width=15, height=10, unit='cm')


ERIN.sig$KEGGPathways<-as.factor(ERIN.sig$"KEGG Pathways")  
ERIN.count<- ERIN.sig%>% 
  select(FeatureID, KEGGPathways, Comment) %>% 
  separate_rows(KEGGPathways, sep = ';') %>% 
  group_by(KEGGPathways, Comment ) %>% 
  count(KEGGPathways) %>%
   drop_na(KEGGPathways)%>% filter(Comment != "Only present in control")
   ERIN.count <- within(ERIN.count, n[Comment=="Downregulated"] <- (n*(-1))) 
class.plot<- ggplot(ERIN.count)+
geom_bar(aes(y=KEGGPathways, x=n,fill= Comment), stat="identity", alpha=0.7)+
   scale_fill_manual(values = c("lightsteelblue","tomato2"))+
  theme_classic()+xlab("Number of Features")+ylab("KEGG Pathways")+
  labs(title="Up and down regulated KEGG pathways of Isolate 1177")
class.plot

figure_file <- file.path(figures_dir, 'KEGGrank_ERIN.png')
ggsave(figure_file, class.plot, width=15, height=10, unit='cm')


ERLE.sig$KEGGPathways<-as.factor(ERLE.sig$"KEGG Pathways")  
ERLE.count<- ERLE.sig%>% 
  select(FeatureID, KEGGPathways, Comment) %>% 
  separate_rows(KEGGPathways, sep = ';') %>% 
  group_by(KEGGPathways, Comment ) %>% 
  count(KEGGPathways) %>%
   drop_na(KEGGPathways)%>% filter(Comment != "Only present in control")
    ERLE.count <- within(ERLE.count, n[Comment=="Downregulated"] <- (n*(-1)))
class.plot<- ggplot(ERLE.count)+
geom_bar(aes(y=KEGGPathways, x=n,fill= Comment), stat="identity", alpha=0.7)+
   scale_fill_manual(values = c("lightsteelblue","tomato2"))+
  theme_classic()+xlab("Number of Features")+ylab("KEGG Pathways")+
  labs(title="Up and down regulated KEGG pathways of Isolate 1154")
class.plot

figure_file <- file.path(figures_dir, 'KEGGrank_ERLE.png')
ggsave(figure_file, class.plot, width=15, height=10, unit='cm')


ERLE.count$isolate<- 1154
ERIN.count$isolate<- 1177
BOCU.count$isolate<- 1246
class.rank.all<- rbind(ERLE.count, ERIN.count, BOCU.count)

class.plot<- ggplot(class.rank.all)+
geom_bar(aes(y=KEGGPathways, x=n,fill= Comment), stat="identity", alpha=0.7)+
  facet_wrap(~isolate)+
  scale_fill_manual(values = c("lightsteelblue","tomato2"))+
  theme_classic()+xlab("Number of Features")+ylab("KEGG Pathways")+
  labs(title="Up and down regulated KEGG pathways")
class.plot

figure_file <- file.path(figures_dir, 'KEGG_rank_all.png')
ggsave(figure_file, class.plot, width=30, height=10, unit='cm')

```




## 4.38 Upregulated Secondary Metabolites
```{r}

BOCU.sig$KEGGPathways<-as.factor(BOCU.sig$"KEGG Pathways")  
BOCU.reg<- BOCU.sig %>% 
  #select(FeatureID, KEGGPathways, Comment) %>% 
  separate_rows(KEGGPathways, sep = ';') %>% 
  filter(KEGGPathways== 'Biosynthesis of secondary metabolites')%>% 
 filter(Comment == "Upregulated")

table_file <- file.path(tables_dir, 'Upreg.SecondaryMetabolites_BOCU.csv')
write_csv( BOCU.reg, table_file)
  
ERIN.sig$KEGGPathways<-as.factor(ERIN.sig$"KEGG Pathways")  
ERIN.reg<- ERIN.sig%>% 
  #select(FeatureID, KEGGPathways, Comment) %>% 
  separate_rows(KEGGPathways, sep = ';') %>% 
  filter(KEGGPathways== 'Biosynthesis of secondary metabolites')%>% 
 filter(Comment == "Upregulated")
table_file <- file.path(tables_dir, 'Upreg.SecondaryMetabolites_ERIN.csv')
write_csv( ERIN.reg, table_file)

ERLE.sig$KEGGPathways<-as.factor(ERLE.sig$"KEGG Pathways")  
ERLE.reg<- ERLE.sig%>% 
  #select(FeatureID, KEGGPathways, Comment) %>% 
  separate_rows(KEGGPathways, sep = ';') %>% 
  filter(KEGGPathways== 'Biosynthesis of secondary metabolites')%>% 
 filter(Comment == "Upregulated")
table_file <- file.path(tables_dir, 'Upreg.SecondaryMetabolites_ERLE.csv')
write_csv( ERLE.reg, table_file)
```

### All dysregulated features

```{r}
# Create an AUC matrix that contains only features that are differentially expressed

sig_features.matrix <- norm.matrix[row.names(norm.matrix) %in% sig_features, ]

row.names(sig_features.matrix) <- compounds_table$name4plot[match(row.names(sig_features.matrix), compounds_table$FeatureID)]

```





