---
title: "Caffeic Acid Germination"
author: "Taylor A Portman"
date: "2022-09-26"
output: html_document
---
Caffeic Acid Germination Experiment

Objective:
Investigate the role of CA on the germination of native and invasive grasses common to the SRER.

Hypothesis:
Caffeic acid produced by endophytic fungi associated with the invasive grass E. lehmanniana decreases the germination of native grasses while increasing the germination of the invasive grass.

Plant seeds used: Eragrostis lehmanniana (ERLE), Eragrostis intermedia (ERIN), Bouteloua curtipendula (BOCU), Digitaria california (DICA), Leptachola dubia (LEDU)

10 surface sterilized seeds of each grass species were  grown in three media treatments: control MEA media, 0.5mM Caffeic acid amended MEA media, and fungal lawns of Fusarium sp. isolate 1154 on MEA media. 

```{r}
library(tidyverse)
library(ggplot2)
getwd()
setwd("/Volumes/GoogleDrive/My Drive/projects/MastersThesis/R/ERLE_fungal_symbionts")
dat<-read.csv("data_raw/CaffeicAcid_SeedGermination.csv")

data_clean = '/Volumes/GoogleDrive/My Drive/projects/MastersThesis/R/ERLE_fungal_symbionts/data_clean'

figures= '/Volumes/GoogleDrive/My Drive/projects/MastersThesis/R/ERLE_fungal_symbionts/CaffeicAcidGermination/figures'
```

```{r}
dat$Germ<-dat$Day14.30Sep
germ<-dat%>%
  dplyr::select( Sample, Seed, Media, Germ)

head(germ)

means.grass <- germ %>% 
  dplyr::group_by(Seed, .add=TRUE) %>%
  dplyr::summarise(
    means=mean(Germ))

germ.dat<-germ%>%
  dplyr::left_join(means.grass, by="Seed", copy=FALSE)

germ.dat$average<- germ.dat$Germ/ germ.dat$means
head(germ.dat)
```

Save Clean Data
```{r}
fileName = paste(data_clean, 'CaffeicAcidGermination_Clean.csv',sep = '/')
write.csv(germ.dat2, fileName )


```


Plot
```{r}
ggplot(germ.dat, aes(Seed, Germ, fill=Media))+
  geom_boxplot()

germ.dat2<-germ.dat%>%
  dplyr::filter(Seed != "CTRL", Seed != "DICA")

ggplot(germ.dat2, aes(Seed, Germ, fill=Media))+
  geom_boxplot()

```


