---
title: "Compound Discover Data Cleaning "
output: html_notebook
---
---
title: "CompoundDiscover data clean+combine All Isolates+ seeds"
output: html_notebook
---

# 1. Importing Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(ggpubr)
library(ggsci)
library(ggpolypath)
library(venn)
library(dplyr)
source('functions_cdis_exploration.R')
```


Change based on what data you are using
```{r}
data= 'RP'
```

# 2. Create directory

Import data from compound discover.

```{r set_path, message=FALSE}
#set path variables
project_dir <- setwd("/Volumes/GoogleDrive/My Drive/projects/ERLE_fungal_symbionts/IsolatesAll")

# Give a name for your project so the results are all grouped together in a directory with the same name
project_name <- 'RP_All_checked'
figures_dir <- file.path(project_dir, paste0(project_name, '_output_figures'))
tables_dir <- file.path(project_dir,  paste0(project_name, '_output_tables'))

# Create output directories
dir.create(figures_dir, showWarnings = FALSE)

dir.create(tables_dir, showWarnings = FALSE)
```

## 3. Import Data

```{r import_data, message=FALSE}
# Import data tables
cd_results_file <- file.path(project_dir, 'Data', 'CompoundsChecked_Again.csv')
cd_results_table <- read_csv(cd_results_file)

cd_results_table$DeltaMass<- as.numeric(cd_results_table$"Annot. DeltaMass [ppm]")
cd_results_table$MW<- as.numeric(cd_results_table$"Calc. MW")

# Select columns needed for downstream analysis
#Cleaned for Delta mass ppm ..5--.5 and duplicate molecular weights

#drop.cols<- c(' Area: 23')
#cd_results_table<-dplyr::select(cd_results_table, !drop.cols)
cd_results_table <- cd_results_table %>% filter(Checked!="FALSE")
  
  
 # summarize_at(vars(matches("Group Area: ")), mean,.drop=FALSE)%>%
  
cd_results_table<- cd_results_table%>%
  mutate(FeatureID = paste0('Feature',formatC(n():0001, width = 4, flag = '0'))) %>%
 dplyr:: select(FeatureID, Name, Formula, MW, contains('AnnotSource:'), contains('Results'), contains('Pathways'), 
         contains('Group Area:'), contains('Gap Status:')) %>% 
   
  
  # Differentiate between features that share the same name using "peak#" at the end of the name
  group_by(Name) %>%
  add_count(Name) %>% 
  
  # Create variable with names for plotting (useful in following scripts)
  mutate(name4plot = ifelse(is.na(Name), FeatureID, ifelse(n == 1, Name, paste0(Name, '-peak', n():1)))) %>% 
 dplyr:: select(!n) %>% 
  ungroup()

```


## 4. Import and combine metadata
```{r}

# Import metadata and fix names
metadata_file <- file.path(project_dir, 'data', 'metadata.csv')
metadata <- read_csv(metadata_file)

#metadata$`Sample Identifier` <- str_remove(metadata$`Sample Identifier`, 'Saleska_')
#metadata$`Sample Identifier` <- str_remove(metadata$`Sample Identifier`, '_23.*')

# Select only the useful columns and fix column names

metadata <- metadata %>% 
  dplyr::select(SampleID, Plant, Fungi) 
  #rename(metadata,SampleID = `Sample Identifier`)

table_file <- file.path(tables_dir, 'fixed_metadata.csv')
write_csv(metadata, table_file)


```
# 5. Data Manipulation and thermodynamic indices calculations

```{r data, message=FALSE, warning=FALSE}

# Split formula column into elemental counts
cd_results_table <- separate_formula(cd_results_table)

# Calculate ratios and thermodynamic indices
cd_results_table <- calc_ratios_n_idxs(cd_results_table)

# Calculate classes
cd_results_table <- calc_classes(cd_results_table)

table_file <- file.path(tables_dir, 'RPcompounds.table.csv')
write_csv(cd_results_table, table_file)

# Gather area under the curve (AUC) values per sample

compounds_table <- cd_results_table %>% 
  select(-contains('Gap Status:')) %>% 
  
  gather(contains('Area:'), key = 'SampleID', value = 'AUC') %>% 
  filter(AUC > 0)
compounds_table$SampleID <- str_remove(compounds_table$SampleID, 'Group Area: ')

compounds_table$Class <- gsub( "Lignin", "Phenol", compounds_table$Class)
compounds_table$Class <- gsub("Tannin","Other",  compounds_table$Class)


# Save gap-filled table to be used in Statistical Analysis

table_file <- file.path(tables_dir, 'gap_filled_compounds_table.csv')
write_csv(compounds_table, table_file)

```
### Gap filtering (only for unlabel data)

```{r}
# Gather "Gap Status" for filtering 

  gap_status <- cd_results_table %>% 
    select(FeatureID, contains('Gap Status:')) %>% 
    gather(contains('Gap Status:'), key = 'SampleID', value = 'gap_status')
  gap_status$SampleID <- str_remove(gap_status$SampleID, 'Gap Status: ')
  
  ## Filtering
  
  compounds_table <- left_join(compounds_table, gap_status, by = c('FeatureID', 'SampleID')) 
  


```

``` {r}
# Add metadata information
compounds_table$SampleID<- as.factor(compounds_table$SampleID)
metadata$SampleID<- as.factor(metadata$SampleID)
compounds_table <-  left_join(compounds_table, metadata, by = 'SampleID')

table_file <- file.path(tables_dir, 'compounds_table.csv')
write_csv(compounds_table, table_file)

```


Doesn't work because using already gap filled data: would need gaps
```{r venn_time, message=FALSE}
compounds_table_nogap<- compounds_table %>%
filter(gap_status !="Full gap")

# Venn diagram of the compounds present in each isolate
CTRL_list <- compounds_table_nogap %>% 
  filter(Fungi == 'CTRL') %>% 
  select(FeatureID) %>% 
  distinct()

F1177_list <- compounds_table_nogap %>% 
  filter(Fungi == '1177') %>% 
  select(FeatureID) %>% 
  distinct()

F1246_list <- compounds_table_nogap %>% 
  filter(Fungi == '1246') %>% 
  select(FeatureID) %>% 
  distinct()

F1154_list <- compounds_table_nogap %>% 
  filter(Fungi == '1154') %>% 
  select(FeatureID) %>% 
  distinct()

my_list <- list(CTRL = CTRL_list$FeatureID,
                F1177 = F1177_list$FeatureID,
                F1246 = F1246_list$FeatureID,
                F1154 = F1154_list$FeatureID)

my_colors <- c('grey', 'blue', 'green', 'pink')

figure_file <- file.path(figures_dir, 'venn_isolates.png')
png(figure_file)
venn_time <- plot_venn(my_list, my_colors)
dev.off()

venn_time <- plot_venn(my_list, my_colors)
```
``` {r}
# Add metadata information
compounds_table_nogap$SampleID<- as.factor(compounds_table_nogap$SampleID)
metadata$SampleID<- as.factor(metadata$SampleID)
compounds_table <-  left_join(compounds_table_nogap, metadata, by = 'SampleID')

table_file <- file.path(tables_dir, 'compounds_table_nogap.csv')
write_csv(compounds_table, table_file)

```


# 6. Van Krevelin
```{r vank_material, message=FALSE}

# Ven Krevelen Diagram based on type of organic material
vank_material <- plot_vank(compounds_table, Class)
vank_material

figure_file <- file.path(figures_dir, 'Venk_diagram_all.jpg')
ggsave(figure_file, vank_material, width=30, height=20, unit= 'cm')

```

```{r}
##Number of unique compounds (some Names/ compounds are replicated)
distinct<- distinct(cd_results_table, Name, .keep_all=TRUE)
#unique<- unique(cd_results_table$Name)

##414Unique compounds
```